---
title: "Import PyPSA-Earth model dataset"
subtitle: "build Switch and energyRt models, update PyPSA-Earth input"
author: "Multi-Country Electricity Transition Potential and Challenges"
date:  "`r Sys.Date()`"
format: 
  # html:
  #   code-fold: true
  #   fig-width: 8
  #   fig-height: 6
  #   toc: true
  pdf:
    fig-width: 7
    fig-height: 5  
    toc: true
execute:
  echo: false # should the code be echoed in the output document?
  eval: true # should the code be evaluated? (doesn't override chunk settings)
  error: false # Include errors in the output (false == stop on error)
  # warning: false # Include warnings in the output
  # message: false # Include messages in the output

# Model parameters
params:
  # iso3c: BGD # country code
  # iso3c: CHL
  # iso3c: CHN
  # iso3c: COL
  # iso3c: IND 
  # iso3c: KAZ
  iso3c: MYS
  # iso3c: THA
  # iso3c: VNM
  # Model choice, defined by selection of a the YAML file
  yaml_path: "config/mod_toy" # file: "{yaml_path}/config_{iso3c}.yml
  # yaml_path: "config/mod_adm" # file: "{yaml_path}/config_{iso3c}.yml
  # yaml_path: "config/mod_custom" # file: "{yaml_path}/config_{iso3c}.yml

  # Execution controls
  save_raw_csv: false # save imported PyPSA-Earth data as csv files
  build_switch: true # write/overwrite Switch model input files
  solve_switch: false # solve Switch model
  build_energyRt: true # write/overwrite energyRt model input files
  solve_energyRt: false # solve energyRt model

  force: true # try to auto-resolve some declaration errors in data?
  
  # Adjustments to the imported PyPSA-Earth data
  ## 1. Use tech-data from PyPSA-techs dataset (costs, fuels, lifetime)
  use_tech_database: false # 
  ## 2. Capacity factors from MERRA2 (wind and solar only)
  use_merra_data: true # Update capacity factors with (pre-saved) MERRA-2
  weather_year: 2013 # weather year
  sol_cl_tol: 1 # SD loss tolerance [0-1] (1 == 100%, one cluster per region)
  win_cl_tol: 1 # SD loss tolerance [0-1] (1 == 100%, one cluster per region)
  sol_cl_max: 100 # import first X clusters per region
  win_cl_max: 100 # import first X clusters per region
  sol_landuse_up: .1 # max share of used land per cluster for solar PVs [0-1]
  win_landuse_up: .5 # max share of used land per cluster for onshore wind [0-1]
  wif_landuse_up: .0 # max share of used area per cluster for offshore wind 
  # add_offshore_wind: false # add offshore wind - not implemented yet
  ## 3. CCS
  add_ccs: false # add ECCG_CCS and ECOA_CCS technologies if not available
  # if use_tech_database is false, use adjustments:
  ccs_capture_rate: 0.9 # parameter from 0 to 1, share of CO2 captured
  ccs_efficiency_factor: 0.9 # efficiency adjustment
  ccs_fixom_factor: 1.5 # multiplier to fixed O&M (inl. annualized capital)
  ccs_varom_factor: 1.5 # multiplyer to variable O&M
  ## 4. Base year, adjust existing capacity and the load
  set_base_year: 2050
  thermal_capacity_factor: .5 # retired share of existing capacity by the BY
  load_growth_factor: 3 # load multiplier
  
  ## 5. Update PyPSA-Earth dataset with MERRA2 data
  update_pypsa_model: false # only for CL1_TOL1 - one cluster per region
  new_pypsa_file_ending: "_merra.nc" # adds to the modified nc-file
  
  # YAML-file parameters (from 'yaml_path:...') can be overwritten here
  # fig_map_width: 6
  # ...
editor_options: 
  chunk_output_type: console
---

# `r countrycode::countrycode(params$iso3c, "iso3c", "country.name")` model

::: {.center data-latex=""}
**\[WORK IN PROGRESS\]**
:::

  \newline This documented is a reproducible example of "translation" of PyPSA-Earth electric power sector models to Switch and energyRt models. The three models have differences in representation of generating technologies, energy storage, and other model elements. Notes along the script with an interim output describe all steps of the process of the "translation", e.g. remapping the data and parameters from PyPSA to Switch and energyRt, and describes the structure of the model and the data. This experimental exercise has a goal to compare and analyze structural differences between the three models.\
\newline The script in this document reads Yaml configuration file for a country which sets input directories (with PyPSA-Earth files) and destination directories for raw `csv` files of the data-sets, Switch and/or energyRt models. There are two ways to run script in the vignette for a country for which PyPSA-Earth dataset is available. First, it can be downloaded to your project's home directory and "knitted" in RStudio (see "Knit" menu for ".rmd" files). In this case edit the vignette's YAML-header if necessary (top of the file). Alternatively, the whole script can be executed from a R-console (see [Get started](https://opendecarbonization.github.io/revaluation/articles/revaluation.html).)

\newpage


```{r, eval=FALSE}
#| eval: false
## To execute this file:
# adjust parameters (in YAML header), save, and run this command, 
# specifying country and the model
# ?quarto::quarto_render # for help
quarto::quarto_render(
  input = "import_pypsa_model.qmd",
  # output_file = "import_pypsa_model_BGD.pdf",
  # output_file = "import_pypsa_model_IND.pdf",
  # output_file = "import_pypsa_model_THA.pdf",
  # output_file = "import_pypsa_model_KAZ.pdf",
  output_file = "import_pypsa_model_MYS.pdf",
  execute_params = list(
    # iso3c = "BGD", 
    # iso3c = "IND",
    # iso3c = "THA",
    # iso3c = "KAZ",
    iso3c = "MYS",
    yaml_path = "config/mod_toy"
    )
  )
# in the case of errors, run the file chunk-by-chunk
```


## Setup

Load required libraries, set country by `iso3c` code for the evaluation, read yaml file, and load maps.

```{r setup}
#| warning: false
#| message: false
#| collapse: true
#| echo: !expr params$echo

source("functions.R")
stopifnot(mcet_check_installations())
library(tidyverse)
library(ggrepel)
library(glue)
library(sf)
library(fst)
library(data.table)
library(revaluation)
library(energyRt)
# library(future)

# Set country
rev_set_country(params$iso3c)
rev_country()

message(c(
  rep("=", 70), "\n",
  "Importing PyPSA-Earth dataset for ", rev_country(), "\n",
  rep("=", 70)
))
Sys.sleep(7)

# read configuration file for the country
p <- rev_read_yaml(path = params$yaml_path, iso3c = params$iso3c)

# import/merge YAML parameters
for (i in 1:length(params)) {p[[names(params)[i]]] <- params[[i]]}

# load maps
(load(p$files$gis))

if (p$use_tech_database) (load("data/repo_pypsa_techs.RData"))

# print(p)
```

## Import PyPSA-Earth dataset

Read PyPSA's `.nc` file with the model input data, group by sets, and return as a list with tables. Every table is matched with known set of parameters (column names) for further use in the "translation" process to other models.

```{r read_nc}
#| warning: false

# read data from pypsa network nc-file (see config_{}.yml file)
nc <- rev_read_pypsa_nc(p$files$pypsa_network)
class(nc) # tables stored in the named list (use `names(nc)` to print names)

# read PyPSA config file
pypsa_config <- yaml::read_yaml(p$files$pypsa_yaml)

# import the model year
p$base_year <- pypsa_config$load_options$prediction_year
if (params$set_base_year > 0) p$base_year <- p$set_base_year
p$weather_year <- pypsa_config$load_options$weather_year
p$discount <- pypsa_config$costs$discountrate

if (is.null(p$discount)) { 
  # fix empty values for Switch import
  p$discount <- 0
}

if (p$use_merra_data) {
  win_cl_mask <- rev_merra_cl_filemask(p$win_cl_tol)
  sol_cl_mask <- rev_merra_cl_filemask(p$sol_cl_tol)
  if (win_cl_mask == sol_cl_mask) {
    tol_sfx <- glue("_{win_cl_mask}")
  } else {
    tol_sfx <- glue("W{win_cl_mask}_S{sol_cl_mask}")
  }
} else {
  tol_sfx <- ""
}
```

## "BASE" scenario settings
Scenario without any policy intervention. It is used as a base for other scenarios.
```{r}
if (is.null(p$scen)) p$scen <- list()
if (is.null(p$scen$BASE)) p$scen$BASE <- list()

# name of the BASE scenario
p$scen$BASE$name <- glue("scen_pypsa_", p$base_year, tol_sfx,
                         ifelse(p$use_tech_database, "_tdb", ""))

# BASE scenario directory
(p$scen$BASE$root <- file.path(p$dir$scen, p$scen$BASE$name))

# base scenario directory - switch files
if (p$build_switch) {
  (p$scen$BASE$switch <- file.path(p$scen$BASE$root, "switch"))
  if (!dir.exists(p$scen$BASE$switch)) {
    message("Creating: ")
    dir.create(p$scen$BASE$switch, recursive = T)
  }
  message("Switch scenario directory: ", p$scen$BASE$switch)
}

# base scenario directory - energyRt files
if (p$build_energyRt) {
  (p$scen$BASE$energyRt <- file.path(p$scen$BASE$root, "energyRt"))
  if (!dir.exists(p$scen$BASE$energyRt)) {
    message("Creating: ")
    dir.create(p$scen$BASE$root, recursive = T)
  }
  message("energyRt scenario directory: ", p$scen$BASE$energyRt)
}
# knitr::opts_chunk$set(eval = F)
```

### Export to .csv

Optional (see `save_raw_csv` parameter in Yaml header of the vignette) saving of the tables as `.csv` files to the given in yaml file directory.

```{r}
#| eval: !expr params$save_raw_csv

# set directory for csv files ('csv_dir') and execute the chunk
if (params$save_raw_csv) {
  csv_dir <- file.path(p$dir$tmp, "pypsa/raw")
  dir.create(csv_dir, showWarnings = F, recursive = T)
  message("Saving raw csv files: ", csv_dir)
  
  for (i in 1:length(nc)) {
    if (is.data.frame(nc[[i]])) {
      fname <- file.path(csv_dir, paste0(names(nc)[i], ".csv"))
      cat(fname, "\n")
      fwrite(nc[[i]], fname, row.names = F)
    }
  }
}
```


## Regions / buses / load zones

The tree models (PyPSA, Switch, energyRt) have similar notion: buses (PyPSA), load zones (Switch), regions (energyRt). They add spatial dimension to the models, separate modeled objects (generators, loads, etc.) in different zones/regions/buses (used as synonyms in the text below), and link those in the same zones. Though objects in one regions can also be separated (different models use different techniques to do that), they are connected by default. Therefore each particular region/zone/bus can be considered as a "copperplate model" by default (unless they are intentionally disconnected). In contrary, objects in different regions are disconnected by default. Models provide ways to connect them by adding links/transmission/trade and other objects.

Higher number of regions in a model gives better representation of the real-world problem, but it goes with higher computational burden. In this example, the initial model has simplified regions based on scaling down the existing (estimated - see \[...\]) power network. For simplicity we assume that the future demand for electricity will grow ...\[!!!*to discuss*!!!\]... This assumption can be changed later.

Key objects (created in the chunk below):\
`reg_names` - table with names of buses, regions, and load_zones\
`reg_nodes` - `reg_names` with coordinates of network nodes\
`adm1_map` - ggplot map with administrative boundaries (for information and potentially alternative set of scenarios)

```{r}
# names of region (energyRt), buses (PyPSA), and load zones (Switch)
reg_names <- tibble(
  bus = nc$loads_set$loads_t_p_set_i, # PyPSA
  region = rev_bus2reg(nc$loads_set$loads_t_p_set_i, mod = p$model) # energyRt
) |>
  mutate(
    load_zones = region # Switch
  )

# geographic locations of network nodes, one per region
reg_nodes <- nc$buses |> filter(buses_i %in% reg_names$bus) |>
  mutate(
    bus = buses_i,
    lon = buses_x,
    lat = buses_y
  ) |>
  left_join(reg_names, by = "bus") |>
  select(bus, region, load_zones, lon, lat)

reg_nodes_sf <- st_as_sf(reg_nodes, coords = c("lon", "lat")) |>
  st_set_crs(st_crs(gis$mod_map$sf))

# Map with administrative divisions
model_map <- ggplot() + 
  geom_sf(data = gis$mod_map$sf, fill = "wheat") +
  geom_sf(data = reg_nodes_sf, color = "red") +
  # geom_point(aes(lon, lat), data = reg_nodes, col = "red") +
  labs(title = paste0(p$country, " (", nrow(gis$adm1$sf), " administrative regions)"), 
       x = "", y = "") +
  rev_theme_map()

```

### Maps for reports

Creating basic maps (as `ggplot` objects) to use in reports.

```{r}
gis_sf <- gis$mod_map$sf
```

### switch: load_zones.csv

Writing Switch input file with load_zones.

```{r}
#| eval: !expr params$build_switch

dir.create(file.path(p$scen$BASE$switch, "inputs"), showWarnings = F, recursive = T)
load_zones <- tibble(LOAD_ZONE = reg_names$load_zones)
write.csv(load_zones, 
          file = file.path(p$scen$BASE$switch, "inputs/load_zones.csv"), 
          row.names = FALSE, quote = FALSE)
head(load_zones, 3); cat("nrow = ", nrow(load_zones))

stopifnot(!any(is.na((load_zones))))
```

## timeslices / snapshots / timepoints

\[!!ToDo::update!!\] Time dimension in energyRt is set by years and hierarchical sub-annual dimensions (slices). In this project we use use three levels of slices: annual ("ANNUAL"), year-day ("YDAY", from 1 to 365), and hour ("HOUR", 0 to 23). The names and elements of all levels are already saved in `revaluation::rev_data$tsl$d365_h24`. The mapping with PyPSA's "snapshots" are stored in `revaluation::rev_data$timedim`.

```{r}
# timedim <- rev_data$timedim |> mutate(timepoints = slice)
timedim <- revaluation::rev_data$timescales$full_year

# timetable <- revaluation::rev_data$timescales$full_year |> 
#   mutate(
#     d365_h24 = slice,
#     h8760 = paste0("h", formatC(snapshots + 1, width = 4, flag = "0")),
#     timepoints = slice
#   )
# timetable

```

### switch: periods.csv

This table represent time-horizon in Switch.\
The current model has only one (base) year for comparative reasons.

```{r periods.csv}
#| eval: !expr params$build_switch

periods <- tibble(
  INVESTMENT_PERIOD = p$base_year,
  period_start = p$base_year,
  period_end = p$base_year
)
# stop()
stopifnot(!any(is.na(periods)))
write.csv(periods, file = file.path(p$scen$BASE$switch, "inputs/periods.csv"), 
            row.names = FALSE, quote = FALSE)
periods
```

### switch: timeseries.csv

...

```{r}
#| eval: !expr params$build_switch

timeseries <- tibble(
  TIMESERIES = paste0(p$base_year, "_all"),
  ts_period	= p$base_year,
  ts_duration_of_tp = 1L,
  ts_num_tps = 24L * 365L,
  ts_scale_to_period = 1L
)
write.csv(timeseries, file = file.path(p$scen$BASE$switch, "inputs/timeseries.csv"), 
          row.names = FALSE, quote = FALSE)
timeseries
stopifnot(!any(is.na((timeseries))))
```


### switch: timepoints.csv

```{r}
#| eval: !expr params$build_switch

timepoints_map <- tibble(
  year = p$base_year, # assuming one year (temporary solution)
  slice = timedim$timepoints,
  timepoint_id = 
    paste(p$base_year,
      rep(
        timedim$timepoints,
        length(timeseries$TIMESERIES)
      ), sep = "_"
  ),
  timeseries = rep(
    timeseries$TIMESERIES, 
    each = length(timedim$timepoints)
  ),
  timestamp = "."
) |>
  left_join(timedim, by = "slice")
timepoints <- timepoints_map |> select(timepoint_id, timeseries, timestamp)
stopifnot(!any(is.na((timepoints))))
write.csv(timepoints, file = file.path(p$scen$BASE$switch, "inputs/timepoints.csv"),
          row.names = FALSE, quote = FALSE)
head(timepoints)
```

## Commodities / carriers / energy sources

Commodities (energyRt) / carriers (PyPSA) / energy sources\* (Switch)\
\* Switch recognizes **fuels** and **non-fuel energy sources**.

Key objects (created in the chunk below):\
`comm` - a table with names of carriers, commodities

```{r}
# names of commodities
carriers <- nc$carriers |>
  mutate(
    comm = carriers2comm(toupper(carriers_i)),
    drop = duplicated(comm) | is.na(comm)
  )

# drop duplicates, ...
comm <- filter(carriers, !drop) |>
  # assign storage-commodity, slice-level, and supplied (market) commodity
  mutate(
    stg = grepl("HYD|PHS|H2|HGN|BTR", comm), # storage-commodity
    timeframe = if_else(stg, "HOUR", "ANNUAL"), # slice-level
    # limtype = if_else(stg, "", ""), # storage type
    sup = grepl("ANNUAL", timeframe), # supplied commodity
    switch_nonfuel = carriers_co2_emissions == 0
  )
```

### Add CCS-fuels
In the case if CCS technologies are not included, they will be added for ECOA and ECCG with generic assumptions of 90% efficiency, 50% higher variable costs, and 
```{r, eval=params$add_ccs}
#| eval: !expr params$add_ccs

# add COA_CCS and GAS_CCS
comm_ccs <- comm |> 
  filter(grepl("COA|GAS", comm)) |>
  mutate(
    carriers_i = paste(carriers_i, "ccs", sep = "_"),
    carriers_nice_name = paste(carriers_nice_name, "CCS", sep = " "),
    comm = paste(comm, "CCS", sep = "_"),
    carriers_co2_emissions = carriers_co2_emissions * params$ccs_capture_rate
  )

comm <- comm |> bind_rows(comm_ccs) |> unique()
```

### energyRt: commodities

```{r}
#| eval: !expr params$build_energyRt

# create commodities objects
ELC <- newCommodity("ELC", timeframe = "HOUR")
CO2 <- newCommodity("CO2", timeframe = "ANNUAL")
repo_comm <- newRepository("commodities", ELC, CO2)
if (nrow(comm) > 0) {
  for (i in 1:nrow(comm)) {
    com <- newCommodity(
      name = comm$comm[i],
      desc = comm$carriers_nice_name[i],
      timeframe = comm$timeframe[i],
      misc = list(color = comm$carriers_color[i])
    )
    if (comm$carriers_co2_emissions[i] != 0) {
      com <- update(com, 
                    emis = list(comm = "CO2", 
                                emis = comm$carriers_co2_emissions[i]))
    }
    repo_comm <- add(repo_comm, com)
    rm(com)
  }
} 
repo_comm@data |> names()
```


### switch: fuels.csv

Fuels in Switch model are energy sources with associated costs (by year and region, see "fuel_costs" module) and CO2 intensity. They cannot (!!!or can?) be considered as variable an energy source (example: coal, gas, biofuel).

```{r}
#| eval: !expr params$build_switch

fuels <- comm |>
  filter(!switch_nonfuel | grepl("H2|BIO|NUC", comm)) |>
  mutate(
    fuel = comm,
    co2_intensity = carriers_co2_emissions,
    upstream_co2_intensity = 0.
  ) |>
  select(fuel, co2_intensity, upstream_co2_intensity)
stopifnot(!any(is.na((fuels))))
write.csv(fuels, file = file.path(p$scen$BASE$switch, "inputs/fuels.csv"),
          row.names = FALSE, quote = FALSE)
fuels
```


### switch: non_fuel_energy_sources.csv

This type indicates that the energy source can be variable (if declared in "gen_info" module and "variable_capacity_factors"); they don't have associated costs, energy markets. Typical example: wind energy, solar energy, electricity.\
Energy sources with (no costs in the model) and without variable capacity factors can be declared either as fuels or non-fuels. All declarations should be done only once.

```{r}
#| eval: !expr params$build_switch

## switch: non_fuel_energy_sources.csv ####
non_fuel_energy_sources <- comm |>
  filter(!(comm %in% fuels$fuel)) |>
  rename(energy_source = comm) |>
  select(energy_source) |>
  bind_rows(
    tibble(
      energy_source = c("Electricity")
    )) # ??? can it be ELC?

stopifnot(!any(is.na((non_fuel_energy_sources))))
write.csv(non_fuel_energy_sources, 
          file = file.path(p$scen$BASE$switch, "inputs/non_fuel_energy_sources.csv"),
          row.names = FALSE, quote = FALSE)
non_fuel_energy_sources
```

## Supply / markets

Supply (energyRt) / markets (Switch)

```{r}
supp <- comm |> filter(sup)
```

### energyRt: supply

```{r}
#| eval: !expr params$build_energyRt

repo_sup <- newRepository("supply")
for (i in 1:nrow(supp)) {
  sup <- newSupply(
    name = paste0("SUP_", supp$comm[i]),
    commodity = supp$comm[i]
    # desc = supp$carriers_nice_name[i],
  )
  # !!! update costs?
  repo_sup <- add(repo_sup, sup)
  rm(sup)
}
repo_sup@data |> names()
```

### switch: fuel_cost.csv

```{r}
#| eval: !expr params$build_switch

fuel_cost <- expand_grid(
  load_zone	= reg_names$region, 
  fuel = fuels$fuel, # ??? can non-fuels be here?
  period = periods$INVESTMENT_PERIOD,
) |>
mutate(
  fuel_cost = 0. # ??? No data in nc.file. Included in marginal costs?
)

write.csv(fuel_cost, 
          file = file.path(p$scen$BASE$switch, "inputs/fuel_cost.csv"),
          row.names = FALSE, quote = FALSE)
cat("file: inputs/fuel_cost.csv")
fuel_cost
stopifnot(!any(is.na(fuel_cost)))
```

## Demand / load profiles

Demand (energyRt) / loads (PyPSA & Switch) profiles
```{r}
dem <- nc$loads |>
  left_join(reg_names, by = c(loads_t_p_set_i = "bus")) |>
  left_join(timedim, by = "snapshots") |>
  mutate(
    yday = as.integer(str_extract(slice, "[0-9]++")),
    hour = as.integer(str_extract(slice, "[0-2][0-9]$")),
    loads_t_p_set = params$load_growth_factor * loads_t_p_set
    )

loads_byday <- dem |>
  group_by(hour, region) |>
  summarise(
    load_min = min(loads_t_p_set),
    load_max = max(loads_t_p_set),
    load_mean = mean(loads_t_p_set),
    .groups = "drop"
  )

fig_loads_hour <- ggplot(dem) +
  geom_line(aes(x = hour, y = loads_t_p_set/1e3, color = yday, group = yday),
            alpha = .5) +
  # geom_ribbon(aes(hour, ymin = load_min, ymax = load_max)) +
  scale_color_viridis_c(option = "H", name = "yday") +
  facet_wrap(~ region, ncol = 1, scales = "free_y", strip.position = "right") + 
  labs(x = "hour", y = "GWh", 
       title = "Hourly load profile by day of the year") +
  rev_theme_map()

fig_loads_yday <- ggplot(dem) +
  geom_line(aes(x = yday, y = loads_t_p_set/1e3, color = hour, group = hour),
            alpha = .5) +
  # geom_ribbon(aes(hour, ymin = load_min, ymax = load_max)) +
  scale_color_viridis_c(option = "D", name = "hour", limits = c(0, 23)) +
  facet_wrap(~ region, ncol = 1, scales = "free_y", strip.position = "right") + 
  labs(x = "day of year (yday)", y = "GWh", 
       title = "Load profile by day of year and hour") +
  rev_theme_map()
```

```{r}
#| fig.height: 8
#| echo: false
fig_loads_hour
```

```{r}
load_by_reg <- dem |>
  group_by(region) |>
  summarise(
    GWh = sum(loads_t_p_set, na.rm = T) / 1e3,
    .groups = "drop"
  )

load_by_reg_sf <- gis_sf |>
  left_join(load_by_reg, by = "region")
  
fig_load_by_reg_sf <- ggplot(load_by_reg_sf) +
  geom_sf(aes(fill = GWh)) +
  scale_fill_viridis_c(option = "B") +
  labs(title = "Annual load by region/cluster") +
  rev_theme_map()
```

```{r}
#| echo: false
fig_load_by_reg_sf
```

### energyRt: demand

```{r}
DEMELC <- newDemand(
  name = "DEMELC",
  desc = "loads_t_p_set",
  commodity = "ELC",
  unit = "MWh",
  dem = list(
    region = dem$region,
    slice = dem$timepoints,
    dem = dem$loads_t_p_set
  )
)
```

### switch: loads.csv

```{r}
#| eval: !expr params$build_switch

loads <- dem |>
  left_join(
    select(timepoints_map, slice, timepoint_id),
    by = c(timepoints = "slice")) |>
  mutate(
    LOAD_ZONE = region,
    TIMEPOINT = timepoint_id,
    zone_demand_mw = loads_t_p_set
  ) |>
  select(LOAD_ZONE, TIMEPOINT, zone_demand_mw)
stopifnot(!any(is.na((loads))))
write.csv(loads, file = file.path(p$scen$BASE$switch, "inputs/loads.csv"), 
          row.names = FALSE, quote = FALSE)  
loads
```

## Power network

```{r}
network <- nc$lines |> 
  left_join(reg_nodes, by = c("lines_bus0" = "bus")) |>
  left_join(reg_nodes, by = c("lines_bus1" = "bus")) |>
  mutate(
    losses = round(0.05 * lines_length / 1000, 3),
    trd_name = paste("TRL", region.x, region.y, sep = "_")
    ) # assuming 5% losses per 1000 km

net_map <- model_map +
  geom_segment(aes(x = lon.x, y = lat.x, xend = lon.y, yend = lat.y), 
               data = network, color = "dodgerblue", linewidth = 2) +
  geom_point(aes(lon, lat), data = reg_nodes, col = "red") +
  geom_label_repel(aes(lon, lat, label = bus), data = reg_nodes, alpha = 0.7)

```

### energyRt: trade

```{r}
#| eval: !expr params$build_energyRt

repo_network <- newRepository("network")
if (nrow(network) > 0) {
  for (i in 1:nrow(network)) {
    trd <- newTrade(
      name = network$trd_name[i],
      desc = network$lines_type[i],
      commodity = "ELC",
      routes = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i])
      ),
      trade = data.frame(
        src = c(network$region.x[i], network$region.y[i]),
        dst = c(network$region.y[i], network$region.x[i]),
        teff = rep(1 - network$losses[i], 2)
      ),
      capacityVariable = T,
      invcost = data.frame(
        # costs can be assigned to one of the connected region or both
        # here we split the costs, 50% for each region
        region = c(network$region.x[i], network$region.y[i]),
        invcost = rep(network$lines_capital_cost[i] / 2, 2) * 2 # olife == 2
      ),
      olife = list(olife = 2), # doubled annualized invcost for consistency 
      capacity = data.frame(
        # year = 
        stock = network$lines_s_nom[i]
      ),
      cap2act = 24*365
    )
    repo_network <- add(repo_network, trd)
    rm(trd)
  }
}
names(repo_network@data)
```

### switch: transmission_lines.csv

Switch model considers transmission lines bi-directional. Key parameters:

```{r}
#| eval: !expr params$build_switch

transmission_lines <- tibble(
  TRANSMISSION_LINE = network$trd_name,
  trans_lz1 = network$region.x,
  trans_lz2 = network$region.y,
  trans_length_km = network$lines_length,
  trans_efficiency = 1 - network$losses,
  existing_trans_cap = network$lines_s_nom,
  trans_new_build_allowed = network$lines_s_nom_extendable
)
transmission_lines

write.csv(transmission_lines, 
          file = file.path(p$scen$BASE$switch, "inputs/transmission_lines.csv"), 
          row.names = FALSE, quote = FALSE)  
stopifnot(!any(is.na((transmission_lines))))
```

```{r}
#| echo: false
net_map

# ggsave("tmp/network.png", net_map, scale = 2, width = 3.5, height = 4)
```

## Solar capacity factors

Weather factors (energyRt) / renewable profiles (PyPSA) / variable_capacity_factors (Switch).

```{r}
solar <- nc$generators_cf |>
  # filter(grepl("[0-9] solar$", generators_t_p_max_pu_i)) |>
  filter(grepl(" solar$", generators_t_p_max_pu_i)) |>
  rename(cf = generators_t_p_max_pu) |>
  mutate( # !!! ToDo: write functions !!!
    # bus = str_extract(generators_t_p_max_pu_i, "[A-Z]+.[0-9]+"),
    bus = str_replace(generators_t_p_max_pu_i, " solar$", ""),
    tech_pypsa = str_trim(str_replace(generators_t_p_max_pu_i, bus, "")),
    tech = "ESOL", # assuming one technology per region
    weather = "WSOL"
    ) |>
  left_join(reg_names, by = "bus") |>
  left_join(timedim, by = "snapshots") |>
  as.data.table()

  if (is.null(solar$offshore)) {
    solar[, offshore := FALSE]
  }
```

### (Optional) replace with MERRA2 data

```{r}
#| eval: !expr params$use_merra_data

if (params$use_merra_data) {
  # win_cl_mask <- rev_merra_cl_filemask(p$win_cl_tol)
  # # sol_cl_mask <- rev_merra_cl_filemask(p$sol_cl_tol)
  p$files$merra_sol <- 
    fp(p$dir$data_mod, glue("merra_sol_{sol_cl_mask}_{p$weather_year}.fst"))

  cf_sol <- read_fst(p$files$merra_sol, as.data.table = T) |>
    filter(cluster <= p$sol_cl_max) |> as.data.table()
    
  # (load(file.path(p$dir$data, "merra", "sol_sf.RData")))
  unique(cf_sol$cluster)
  
  # drop offshore data
  if (is.null(cf_sol$offshore)) {
    cf_sol[, offshore := F]
  }

  sol_cf <- solar |> 
    left_join(cf_sol, by = c("region" = "reg_off", "offshore", "slice")) |>
    # left_join(filter(cf_win, !offshore)) |>
    filter(!is.na(offshore)) |> # drop regions with no data (low potential)
    select(-cf) |>
    rename(cf = scf_tl) |>
    select(-starts_with("scf_"), -year) |>
    mutate(
      tech = make_tech_name(tech, as.integer(cluster)),
      weather = make_tech_name(weather, as.integer(cluster))
    )
  
  # 
  pypsa_update_solar <- sol_cf |> 
    replace_na(list(cf = 0)) |>
    select(generators_t_p_max_pu_i, snapshots, cf) |>
    rename(generators_t_p_max_pu = cf)  
  
  sol_cf$cf |> summary()
  sol_cf$offshore |> summary()
  solar <- sol_cf
}

```

\[ToDo:: make by cluster\]

```{r}
#| echo: false

# if (is.null(solar$offshore)) solar$offshore <- FALSE
solar_cf <- solar |>
  group_by(across(any_of(c("region", "reg_off", "offshore")))) |>
  summarise(
    cf = mean(cf, na.rm = T), 
    .groups = "drop"
  )

solar_cf_sf <- gis_sf |>
  full_join(
    solar_cf
    # by = "region", 
    # relationship = "many-to-many"
  )

fig_solar_cf_sf <- ggplot(solar_cf_sf) +
  geom_sf(aes(fill = cf)) +
  scale_fill_viridis_c(name = "CF", option = "C") +
  labs(title = "Solar annual average capacity factor") +
  rev_theme_map()

fig_solar_cf_sf
```

### energyRt: weather factors (solar)

```{r}
#| eval: !expr params$build_energyRt

repo_solar_cf <- newRepository("repo_solar_cf")
if (nrow(solar) > 0) {
  for (w in unique(solar$weather)) {
    x <- filter(solar, weather %in% w, !offshore)
    WSOL <- newWeather(
      name = w,
      desc = "Solar generation profile",
      timeframe = "HOUR",
      weather = data.frame(
        region = x$region,
        slice = x$timepoints,
        wval = x$cf
      )
    )
    repo_solar_cf <- add(repo_solar_cf, WSOL); rm(WSOL)
  }
}
repo_solar_cf@data |> names()
```

### switch: variable_capacity_factors (solar)

```{r}
#| eval: !expr params$build_switch

variable_capacity_factors_ESOL <- solar |>
  left_join(timepoints_map, by = c("timepoints" = "slice")) |>
  rename(
    timepoint = timepoint_id,
    gen_max_capacity_factor = cf
  ) |>
  mutate(
    GENERATION_PROJECT = paste0(region, "_", tech)
  ) |>
  select(GENERATION_PROJECT, timepoint, gen_max_capacity_factor)
variable_capacity_factors_ESOL
```

## Onshore wind capacity factors

```{r}
onwind <- nc$generators_cf |>
  filter(grepl("onwind$", generators_t_p_max_pu_i)) |>
  rename(wind_cf = generators_t_p_max_pu) |>
  mutate( # !!! ToDo: write functions !!!
    # bus = str_extract(generators_t_p_max_pu_i, "[A-Z]+.[0-9]+"),
    bus = str_replace(generators_t_p_max_pu_i, " onwind$", ""),
    tech_pypsa = str_trim(str_replace(generators_t_p_max_pu_i, bus, "")),
    tech = "EWIN", # assuming one technology per region
    weather = "WWIN") |>
  left_join(reg_names, by = "bus") |>
  left_join(timedim, by = "snapshots") |>
  mutate(offshore = FALSE)

```

### (Optional) replace with MERRA2 data

```{r}
#| eval: !expr params$use_merra_data
win_cl_mask <- rev_merra_cl_filemask(p$win_cl_tol)
p$files$merra_win <- 
  fp(p$dir$data_mod, glue("merra_win_{win_cl_mask}_{p$weather_year}.fst"))

cf_win <- read_fst(p$files$merra_win, as.data.table = T) |>
  filter(cluster <= p$win_cl_max) |> as.data.table()

if (is.null(cf_win$offshore)) cf_win[, offshore := grepl("_off$", reg_off)]
summary(cf_win$offshore) # check if there is offshore data
# if (!params$add_wind_offshore) 
cf_wif <- cf_win[offshore == T] # separate offshore data
cf_win <- cf_win[offshore == F]

onwind_merra <- onwind |> 
  mutate(offshore = FALSE) |>
  left_join(cf_win, by = c("region" = "reg_off", "slice", "offshore")) |>
  # filter(!offshore) |>
  select(-wind_cf) |>
  rename(wind_cf = wcf100m) |> #!!! rename wind_cf and cf_win
  select(-starts_with("wcf"), -year) 

onwind_merra$wind_cf |> summary()
# onwind_merra$offshore |> summary()
onwind_merra$region |> unique()
onwind_merra$generators_t_p_max_pu_i |> unique()

if (p$update_pypsa_model) {
  # option #1 - replace NA's with zeros 
  pypsa_update_onwind <- onwind_merra |> 
    replace_na(list(wind_cf = 0)) |>
    select(generators_t_p_max_pu_i, snapshots, wind_cf) |>
    rename(generators_t_p_max_pu = wind_cf)

  if (F) {# option #2 - keep original PyPSA data for missing techs (NA's)
    onwind_merra |> 
      select(generators_t_p_max_pu_i, snapshots, wind_cf) |>
      rows_patch(
        select(onwind, generators_t_p_max_pu_i, snapshots, wind_cf), 
        by = c("generators_t_p_max_pu_i", "snapshots")
      )
  }
}

# overwrite
onwind <- onwind_merra |>
  filter(!is.na(cluster)) |> # drop regions w/o data (low potential)
  mutate(
    tech = make_tech_name(tech, as.integer(cluster)),
    weather = make_tech_name(weather, as.integer(cluster))
  )  

```

!!! - make by clusters

```{r}
#| ehco: false

# if (is.null(onwind$offshore)) onwind$offshore <- FALSE
onwind_cf <- onwind |>
  group_by(region, offshore) |>
  summarise(
    cf = mean(wind_cf, na.rm = T), 
    .groups = "drop"
  )
onwind_cf_sf <- gis_sf |>
  full_join(
    onwind_cf,
    # by = "region", 
    relationship = "many-to-many") |>
  mutate(
    reg_off = paste(region, if_else(offshore, "off", ""), sep = "_")
  )

fig_onwind_cf_sf <- ggplot(onwind_cf_sf) +
  geom_sf(aes(fill = cf)) +
  scale_fill_viridis_c(name = "CF", option = "D") +
  labs(title = "Wind annual average capacity factor") +
  rev_theme_map()

fig_onwind_cf_sf
```

### energyRt: weather factors (onshore wind)

```{r}
#| eval: !expr params$build_energyRt

repo_onwind_cf <- newRepository("repo_onwind_cf")
if (nrow(onwind) > 0) {
  for (w in unique(onwind$weather)) {
    x <- filter(onwind, weather %in% w)
    WWIN <- newWeather(
      name = w,
      desc = "Onshore wind generation profile",
      timeframe = "HOUR",
      weather = data.frame(
        region = x$region,
        slice = x$timepoints,
        wval = x$wind_cf
      )
    )
    repo_onwind_cf <- add(repo_onwind_cf, WWIN); rm(WWIN)
  }
}
repo_onwind_cf@data |> names()
```

### switch: variable_capacity_factors (onshore wind)

```{r}
#| eval: !expr params$build_switch

variable_capacity_factors_EWIN <- onwind |>
  left_join(timepoints_map, by = c("timepoints" = "slice")) |>
  rename(
    timepoint = timepoint_id,
    gen_max_capacity_factor = wind_cf
  ) |>
  mutate(
    GENERATION_PROJECT = paste0(region, "_EWIN")
  ) |>
  select(GENERATION_PROJECT, timepoint, gen_max_capacity_factor)
variable_capacity_factors_EWIN
```

## Hydro (ror) capacity factors

```{r}
ror <- nc$generators_cf |>
  # filter(grepl("[0-9] ror$", generators_t_p_max_pu_i)) |>
  filter(grepl(" ror$", generators_t_p_max_pu_i)) |>
  mutate( # !!! ToDo: write functions !!!
    # bus = str_extract(generators_t_p_max_pu_i, "[A-Z]+.[0-9]+"),
    bus = rev_extract_bus(generators_t_p_max_pu_i),
    # tech_pypsa = str_replace(generators_t_p_max_pu_i, " ror$", ""),
    # tech_pypsa = str_trim(str_replace(generators_t_p_max_pu_i, bus, "")),
    tech_pypsa = "ror",
    tech = "EROR", # assuming one technology per region
    weather = "WROR") |>
  left_join(reg_names, by = "bus") |>
  left_join(timedim, by = "snapshots")

unique(ror$generators_t_p_max_pu_i)
unique(ror$bus)
ror_cf_sf <- ror |>
  group_by(region) |>
  summarise(
    cf = mean(generators_t_p_max_pu, na.rm = T), 
    .groups = "drop"
  )
ror_cf_sf <- gis_sf |>
  full_join(
    ror_cf_sf,
    by = "region")

fig_ror_cf_sf <- ggplot(ror_cf_sf) +
  geom_sf(aes(fill = cf)) +
  scale_fill_viridis_c(name = "CF", option = "D") +
  labs(title = "Hydro (ROR) annual average capacity factor") +
  rev_theme_map()

```

```{r}
#| echo: false
fig_ror_cf_sf
```

### energyRt: weather factors (run of river)

```{r}
#| eval: !expr params$build_energyRt

repo_ror_cf <- newRepository("repo_ror_cf")
if (nrow(ror) > 0) {
  for (w in unique(ror$weather)) {
    x <- filter(ror, weather %in% w)
    WROR <- newWeather(
      name = w,
      desc = "Run of river hydro generation profile",
      timeframe = "HOUR",
      weather = data.frame(
        region = x$region,
        slice = x$timepoints,
        wval = x$generators_t_p_max_pu
      )
    )
    repo_ror_cf <- add(repo_ror_cf, WROR); rm(WROR)
  }
}
repo_ror_cf@data |> names()

```

### switch: variable_capacity_factors (run of river)

```{r}
#| eval: !expr params$build_switch

variable_capacity_factors_EROR <- ror |>
  left_join(timepoints_map, by = c("timepoints" = "slice")) |>
  rename(
    timepoint = timepoint_id,
    gen_max_capacity_factor = generators_t_p_max_pu
  ) |>
  mutate(
    GENERATION_PROJECT = paste0(region, "_EROR")
  ) |>
  select(GENERATION_PROJECT, timepoint, gen_max_capacity_factor)
variable_capacity_factors_EROR
```

### switch: variable_capacity_factors.csv

Combine all capacity factors of variable technologies and write the csv-file.

```{r}
#| eval: !expr params$build_switch

variable_capacity_factors <- variable_capacity_factors_ESOL |>
  rbind(variable_capacity_factors_EWIN) |>
  rbind(variable_capacity_factors_EROR)

# checks
variable_capacity_factors |> summary()
variable_capacity_factors$GENERATION_PROJECT |> unique()
str_replace(variable_capacity_factors$GENERATION_PROJECT, 
            "^([A-Z])+(|[0-9])+_", "") |> unique()
# stopifnot(!any(is.na((variable_capacity_factors))  
write.csv(variable_capacity_factors,
          file = file.path(p$scen$BASE$switch, "inputs/variable_capacity_factors.csv"),
          row.names = FALSE, quote = FALSE)
variable_capacity_factors
```

## Hydro inflow

Those time series can be used to control charging process in energyRt in storage or supply type of processes, or in generators in Switch or energyRt.

```{r}
inflow <- nc$storage_inflow |>
  mutate( # !!! ToDO: write functions !!!
    # bus = str_extract(storage_units_t_inflow_i, "[A-Z]+.[0-9]+"),
    bus = str_replace(storage_units_t_inflow_i, " hydro", ""),
    storage = str_trim(str_replace(storage_units_t_inflow_i, bus, "")),
    comm = str_sub(toupper(storage), 1, 3),
    stg = paste0("STG_", comm), # in the case it is used in storage class objects
    sup = paste0("SUP_", comm), # in the case it is used in supply class objects
    weather = paste0("W", comm)  # 
  ) |>
  left_join(reg_names, by = "bus") |>
  left_join(timedim, by = "snapshots")

if (nrow(inflow) > 0) {
  unique(inflow$bus)
  unique(inflow$stg)
  unique(inflow$weather)

  inflow_sf <- inflow |>
    group_by(region) |>
    summarise(
      storage_units_t_inflow = mean(storage_units_t_inflow, na.rm = T), 
      .groups = "drop"
    )
  inflow_sf <- gis_sf |>
    full_join(
      inflow_sf,
      by = "region")
  
  fig_inflow_sf <- ggplot(inflow_sf) +
    geom_sf(aes(fill = storage_units_t_inflow)) +
    scale_fill_viridis_c(name = "", option = "D") +
    labs(title = "Annual storage_units_t_inflow") +
    rev_theme_map()
}
```

```{r}
#| echo: false
try(fig_inflow_sf, silent = T)
```

### energyRt: weather factors (hydro storage inflow)

```{r}
#| eval: !expr params$build_energyRt
repo_inflow <- newRepository("repo_inflow")
if (nrow(inflow) > 0) {
  for (w in unique(inflow$weather)) {
    x <- filter(inflow, weather %in% w)
    WHYD <- newWeather(
      name = w,
      desc = "Hydro inflow profile",
      timeframe = "HOUR",
      weather = data.frame(
        region = x$region,
        slice = x$timepoints,
        wval = x$storage_units_t_inflow
      )
    )
    repo_inflow <- add(repo_inflow, WHYD); rm(WHYD)
  }
}
repo_inflow@data |> names()
```

## Generators

```{r}
gen <- nc$generators |>
  # filter(grepl("[0-9] onwind$", generators_t_p_max_pu_i)) |>
  mutate(
    # bus = str_extract(generators_i, "[A-Z]+.[0-9]+"),
    bus = str_replace(generators_i, " [:alnum:]+$", ""),
    comm = carriers2comm(generators_i),
    gen = str_replace(generators_i, bus, ""),
    gen = str_trim(gen),
    tech_name = substr(paste0("E", toupper(gen)), 1, 4), # !!! ToDo: Write a function
    # allow all techs for investment
    generators_p_nom_extendable = if_else(generators_capital_cost > 0, 1L, 0L),
    is_variable = ifelse(grepl("win|sol|ror", generators_i), TRUE, FALSE) #!!!
  ) |>
  left_join(reg_names, by = "bus")
gen$tech_name[grepl("EONW", gen$tech_name)] <- "EWIN" # !!! ToDo: Write a function
gen$tech_name |> unique()


gen_sf <- gen |>
  group_by(region, tech_name) |>
  summarise(
    GW = sum(generators_p_nom, na.rm = T)/1e3, 
    GW_max = sum(generators_p_nom_max , na.rm = T)/1e3,
    .groups = "drop"
  ) |>
  mutate(offshore = FALSE)

if (params$use_merra_data) {
  (load(file.path(p$dir$data_mod, "win_GW_max_tol99.RData")))
  (load(file.path(p$dir$data_mod, "sol_GW_max_sf_tol99.RData")))
  win_upd <- win_GW_max_sf |>
    st_drop_geometry() |>
    mutate(
      # tech_name = "EWIN",
      tech_name = make_tech_name("EWIN", as.integer(cluster)), 
      # GW = as.numeric(NA),
      .after = "region"
    ) |>
    mutate(GW_max = params$win_landuse_up * win_GW_max) |>
    filter(!offshore) |>
    select(all_of(c("region", "tech_name", "GW_max", "offshore"))) |>
    unique()
    # select(-offshore)
  
  sol_upd <- sol_GW_max_sf |>
    st_drop_geometry() |>
    mutate(
      # tech_name = "ESOL",
      tech_name = make_tech_name("ESOL", as.integer(cluster)), 
      .after = "region"
    ) |>
    filter(!offshore) |>
    mutate(GW_max = params$sol_landuse_up * sol_GW_max) |>
    select(all_of(c("region", "tech_name", "GW_max", "offshore"))) |>
    unique()
    # select(-offshore)

  gen_sf <- gen_sf |>
      rows_update(win_upd, 
                  unmatched = "ignore",
                  by = c("region", "tech_name", "offshore")) |>
      rows_update(sol_upd, 
                  unmatched = "ignore",
                  by = c("region", "tech_name", "offshore"))
}

gen_sf <- gis_sf |>
  full_join(
    gen_sf,
    by = c("region", "offshore"),
    relationship = "many-to-many") |>
  filter(!is.na(tech_name))

fig_gen_cap <- ggplot() +
  geom_sf(data = gis_sf, fill = "wheat") +
  geom_sf(aes(fill = GW), data = select(gen_sf, !offshore)) +
  scale_fill_viridis_c(name = "GW", option = "H", limits = c(0, NA)) +
  labs(title = "Existing (installed) generators, GW") +
  facet_wrap(~tech_name) +
  rev_theme_map() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())

fig_gen_cap_max <- ggplot() +
  geom_sf(data = gis_sf, fill = "wheat") +
  geom_sf(aes(fill = GW_max), data = select(gen_sf, !offshore)) +
  scale_fill_viridis_c(name = "GW", option = "H", limits = c(0, NA)) +
  labs(title = "Maximum capacity by technology, GW") +
  facet_wrap(~tech_name) +
  rev_theme_map() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), 
        axis.text.y = element_blank(), axis.ticks.y = element_blank())

```

```{r}
#| echo: false
fig_gen_cap
fig_gen_cap_max
```

### Add CCS-techs
```{r, eval=params$add_ccs}
#| eval: !expr params$add_ccs
gen_ccs <- gen |> 
  filter(grepl("COA|GAS", comm)) |>
  filter(grepl("ECOA|ECCG", tech_name)) |>
  mutate(
    comm = paste(comm, "CCS", sep = "_"),
    tech_name = paste(tech_name, "CCS", sep = "_"),
    generators_p_nom_extendable = 1L,
    generators_capital_cost = generators_capital_cost * params$ccs_fixom_factor,
    generators_marginal_cost = 
      generators_marginal_cost * params$ccs_varom_factor,
    generators_efficiency = 
      generators_efficiency * params$ccs_efficiency_factor
  )

gen <- gen |> bind_rows(gen_ccs) |> unique()

```

### energyRt: technologies (generators)

```{r}
#| eval: !expr params$build_energyRt

repo_gen <- newRepository("generators")
for (tch in unique(gen$tech_name)) {
  # tch <- unique(gen$tech_name)[1]
  tech <- NULL 
  x <- gen |> filter(tech_name == tch)
  if (p$use_tech_database) {
    # try to import from the database
    tech <- repo_pypsa_techs@data[[tch]] 
    if (!inherits(tech, "technology")) {
      message(tch, 
              " does not exist in `repo_pypsa_techs`, building from scratch")
    }
  }
  if (is.null(tech)) { 
      tech <- newTechnology(
        name = tch,
        desc = x$generators_i[1],
        # desc = paste(x$generators_i, collapse = ", "),
        input = list(comm = unique(x$comm)),
        output = list(comm = "ELC"),
        cap2act = 24*365,
        olife = list(olife = 100), # for 1-year model & annualized costs
        region = unique(x$region),
        ceff = list(
          region = x$region,
          comm = x$comm,
          cinp2use = x$generators_efficiency
        ),
        fixom = list(
          region = x$region,
          fixom = x$generators_capital_cost # annualized
          # invcost = x$generators_capital_cost # overnight
        ),
        varom = list(
          region = x$region,
          varom = x$generators_marginal_cost
        ),
        capacity = list(
          region = x$region,
          stock = x$generators_p_nom
        )
      )
    }
  
  # update pre-existing capacity (from PyPSA-model)
  tech <- update(tech, capacity = data.frame(
    year = p$base_year,
    region = x$region,
    stock = x$generators_p_nom
  ))
  
  # link with capacity factors (solar, wind, hydro)
  # !!! add filter for missing capacity factors
  if (grepl("sol", tch, ignore.case = T)) {
    tech <- update(tech, weather = list(weather = "WSOL", waf.fx = 1))
  } else if (grepl("win", tch, ignore.case = T)) {
    tech <- update(tech, weather = list(weather = "WWIN", waf.fx = 1))
  } else if (grepl("ror", tch, ignore.case = T)) {
    tech <- update(tech, weather = list(weather = "WROR", waf.fx = 1))
  }
  
  # availability of the technology for investment
  extendable <- x$generators_p_nom_extendable
  if (any(extendable != 1)) {
    if (!any(extendable == 1 | extendable == 0)) {
      # error in the data
      print(x)
      stop("Unexpected value in 'generators_p_nom_extendable' {0 or 1}")
    }
   tech <- update(
     tech, 
     end = list(
       region = x$region,
       end = if_else(
         x$generators_p_nom_extendable == 1,
         p$base_year + 50, # open for investment
         p$base_year - 1 # not available for investment
        )
      )
    )
  }
  repo_gen <- add(repo_gen, tech)
}
names(repo_gen@data)
try(draw(repo_gen@data[[1]]))
```

### switch: gen_info (generators)

```{r}
#| eval: !expr params$build_switch

gen_info_tech <- gen |>
  mutate(
    GENERATION_PROJECT = paste0(region, "_", tech_name),
    gen_tech = tech_name,
    gen_load_zone = region,
    gen_connect_cost_per_mw = 0.,
    gen_capacity_limit_mw = generators_p_nom_max, # ".",
    gen_full_load_heat_rate = as.character(
      round(
        convert("GWh/GWh", "MMBtu/MWh", 1 / generators_efficiency), 3)),
    gen_variable_om = generators_marginal_cost,
    gen_max_age = 2, # !!! ToDo: convert to overnight costs
    gen_is_variable = 
      as.integer(
        grepl("ECSP|ESPV|ESOL|EWIN|EWIF|EROR|EHYD", tech_name)), # !!! ToDo: Write function
    gen_full_load_heat_rate = 
      if_else(gen_is_variable == 1, ".", gen_full_load_heat_rate),
    gen_is_baseload = as.integer(!as.logical(gen_is_variable)),
    gen_is_cogen = 0L,
    gen_energy_source	= comm,
    gen_store_to_release_ratio = ".",
    gen_storage_efficiency = ".",
     # gen_unit_size	gen_ccs_capture_efficiency	gen_ccs_energy_load
  )

gen_info_tech$gen_full_load_heat_rate[
  gen_info_tech$gen_is_variable == 1] <- "." 

gen_info_tech$gen_capacity_limit_mw[
  is.infinite(gen_info_tech$gen_capacity_limit_mw)] <- "."

gen_info_csv <- gen_info_tech |>
  select(GENERATION_PROJECT:gen_storage_efficiency)
gen_info_csv$gen_tech |> unique()
stopifnot(!any(is.na((gen_info_csv))))

# drop techs with missing capacity factors

```

### switch: gen_build_predetermined (generators)

```{r}
#| eval: !expr params$build_switch

if (nrow(gen_info_tech) > 0) {
  gen_build_predetermined_tech <- gen_info_tech |>
    filter(generators_p_nom > 0 | generators_p_nom_extendable == 0) |>
    mutate(
      build_year = p$base_year - 1,
      build_gen_predetermined = generators_p_nom
    ) |>
    select(GENERATION_PROJECT, build_year, build_gen_predetermined)

  gen_build_predetermined <- gen_build_predetermined_tech
  # stopifnot(!any(is.na((gen_build_predetermined))))
} else {
  gen_build_predetermined <- NULL
}
```

### switch: gen_build_costs (generators)

```{r}
#| eval: !expr params$build_switch

gen_build_costs <- gen_info_tech |>
  mutate(
    build_year = NA,
    gen_overnight_cost = 0L,
    gen_fixed_om = generators_capital_cost,
    # gen_overnight_cost = generators_capital_cost,
    # gen_fixed_om = 0L,
    gen_storage_energy_overnight_cost = "."
  ) |>
  select(GENERATION_PROJECT, build_year, gen_overnight_cost, gen_fixed_om,
         gen_storage_energy_overnight_cost)

# costs of existing stock
gen_build_costs_0 <- gen_build_costs |>
  filter(GENERATION_PROJECT %in% gen_build_predetermined$GENERATION_PROJECT) |>
  mutate(
    build_year = p$base_year - 1
  )

# investable 
gen_build_costs_1 <- gen_build_costs |>
  filter(
    GENERATION_PROJECT %in% 
      gen_info_tech$GENERATION_PROJECT[
        gen_info_tech$generators_p_nom_extendable == 1]) |>
  mutate(
    build_year = p$base_year
  )
# merge
gen_build_costs <- rbind(gen_build_costs_0, gen_build_costs_1)
rm(gen_build_costs_0, gen_build_costs_1)
gen_build_costs$GENERATION_PROJECT |> unique()

stopifnot(!any(is.na((gen_build_costs))))
head(gen_build_costs)
```

## Storage types

There are several types of storage in the current PyPSA-Earth models:

1.  **accumulators** (of a carrier):

-   data table: `nc$stores`
-   carrier: "battery" or "H2" (electricity or hydrogen)
-   no existing capacity (can be added later)
-   capital costs: `stores_capital_cost`
-   efficiency: 100% (losses are modeled via "links": chargers and dischargers)

2.  **pumped hydro**

-   data table: `nc$storage_units`
-   carrier: "PHS"
-   `storage_units_capital_cost`
-   `storage_units_efficiency_dispatch`
-   `storage_units_efficiency_store` (!!!units? charging efficiency or storing?)

3.  **hydro power plant (PP) with a dam** and *exogenous inflow*:

-   data tables:
    -   `nc$storage_units_set` - sets
    -   `nc$storage_units` - parameters
    -   `nc$storage_inflow` - exogenous inflow factors
-   carrier: "hydro"
-   `storage_units_p_nom` - existing electric capacity
-   `storage_units_max_hours` - storage capacity parameter
-   `storage_units_efficiency_dispatch`\
    ***Note:** the listed types of storage are model/scenario-specific, not generic.*

## Accumulators

Modeled as a generic storage, the model optimizes energy capacity (MWh). Efficiency is assumed 100%, losses are modeled via links ("chargers" and "dischargers" for electricity, "electrolysis" and "fuel cells" for hydrogen).

```{r}
storage <- nc$stores |>
  mutate(
    # bus = str_extract(stores_bus, "[A-Z]+.[0-9]+"),
    bus = str_replace(stores_bus, " [:alnum:]+$", ""),
    comm = carriers2comm(stores_carrier),
    stores = str_replace(stores_i, bus, ""),
    stores = str_trim(stores),
    # stg_name = paste0("STG_", toupper(stores))
    # storage = str_trim(str_replace(storage_units_t_inflow_i, bus, "")),
    comm = str_sub(toupper(stores), 1, 3),
    stg = paste0("STG_", comm),
    comm = str_replace(comm, "^BAT$", "ELC"),
    stg = str_replace(stg, "_BAT", "_BTR")
  ) |>
  left_join(reg_names, by = "bus")
# storage
cat("Found", length(unique(storage$stg)), 
    "stores (accumulators) for carriers:\n"); unique(storage$stores_carrier)
cat("Storage technologies names to create in energyRt:\n"); unique(storage$stg)
```

### energyRt: storage (accumulators)

```{r}
#| eval: !expr params$build_energyRt

repo_stg <- newRepository("storages")
if (nrow(storage) > 0) {
  # s <- unique(storage$stg)[2]
  for (s in unique(storage$stg)) {
    x <- storage |> filter(stg == s)
    stg <- NULL
    if (p$use_tech_database) { # find in database
      stg <- repo_pypsa_techs@data[[s]]
      if (!inherits(stg, "storage")) {
        message(s, 
                " was not found in the `pypsa-technologies` dataset, \n",
                "   building from scratch")
        stg <- NULL
      }
    }
    if (is.null(stg)) {
      stg <- newStorage(
        name = s,
        # desc = paste(x$stores_i, collapse = ", "),
        commodity = unique(x$comm),
        # cap2act = 24*365,
        region = unique(x$region),
        olife = list(olife = 100), # for 1-year model & annualized costs
        fixom = list(
          region = x$region,
          fixom = x$stores_capital_cost # annualized costs
          # invcost = x$stores_capital_cost
        )
      )
    }
    
  # availability of the technology for investment
    extendable <- x$stores_e_nom_extendable
    if (any(extendable != 1)) {
      if (!any(extendable == 1 | extendable == 0)) { # check for errors
        # error in the data
        print(x)
        stop("Unexpected value in 'stores_e_nom_extendable' {0 or 1}")
      }
      stg <- update(
        stg, 
        end = list(
          region = x$region,
          end = if_else(
            x$stores_e_nom_extendable == 1,
            p$base_year + 50, # open for investment
            p$base_year - 1 # not available for investment
          )
        )
      )
    }    
    repo_stg <- add(repo_stg, stg); rm(stg)
  }
}
repo_stg@data |> names()
```

**Note:** Accumulators in Switch are modeled together with "links" (chargers and dischargers) below.

## Pumped hydro (PHS)

This type of storage in PyPSA-Earth has hydro as an `storage_units_carrier`. Though if availability of hydro is not linked with the ability to accumulate (!check!), then the modeling of the carrier can be avoided. The storage itself can be represented as an electricity storage with overall storage capacity equal to `storage_units_p_nom * storage_units_max_hours` in MWh where `storage_units_p_nom` is discharge capacity in MW. This approach is straightforward to reproduce in both Switch and energyRt. A version with hydro carrier can also be modeled, but may require advanced hydro module (will be considered later).

```{r}
unique(nc$storage_units$storage_units_carrier)
phs <- nc$storage_units |>
  filter(grepl("PHS", storage_units_carrier))
if (nrow(phs) > 0) {
  phs <- phs |>
    mutate(
      # bus = str_extract(stores_bus, "[A-Z]+.[0-9]+"),
      bus = storage_units_bus,
      # comm = carriers2comm(storage_units_carrier),
      comm = "ELC", # consider as electricity storage
      stores = str_replace(storage_units_i, bus, ""),
      stores = str_trim(stores),
      stg_name = paste0("STG_", toupper(stores)),
      extendable = if_else(storage_units_capital_cost > 0, 1, 0) #!!! assumed
    ) |>
    left_join(reg_names, by = "bus")
}
```

#### energyRt: storage (PHS)

```{r}
#| eval: !expr params$build_energyRt

repo_phs <- newRepository("phs")
if (nrow(phs) > 0) {
  for (s in unique(phs$stg_name)) {
    x <- phs |> filter(stg_name == s)
    stg <- newStorage(
      name = s,
      # desc = paste(x$stores_i, collapse = ", "),
      commodity = unique(x$comm),
      cap2stg = mean(x$storage_units_max_hours),
      region = unique(x$region),
      seff = list(
        region = x$region,
        inpeff = x$storage_units_efficiency_store,
        outeff = x$storage_units_efficiency_dispatch
      ),
      olife = list(olife = 100), # for annualized costs
      fixom = list(
        region = x$region,
        fixom = x$storage_units_capital_cost # using annualized costs
      ),
      # invcost = list(
      #   region = x$region,
      #   invcost = x$storage_units_capital_cost
      # ),
      capacity = list(
        region = x$region,
        stock = x$storage_units_p_nom
      )
    )
    if (any(x$extendable == 0)) { # availability on the market
      stg <- update(
        stg,
        end = list(
          region = x$region,
          end = if_else(x$extendable == 1, 
                        p$base_year,
                        p$base_year - 1
                        )
        ))
      
    }
    repo_phs <- add(repo_phs, stg); rm(stg)
  }
}
repo_phs@data |> names()
```

#### switch: gen_info (PHS)

Pumped hydro storage, adding to `gen_info` table.

```{r}
#| eval: !expr params$build_switch

if (nrow(phs) > 0) {
  gen_info_phs <- phs |>
    mutate(
      GENERATION_PROJECT = paste0(region, "_", stg_name),
      gen_tech = stg_name,
      gen_load_zone = region,
      gen_connect_cost_per_mw = 0.,
      gen_capacity_limit_mw = storage_units_p_nom, # ".",
      gen_full_load_heat_rate = ".",
      gen_variable_om = 0.,
      gen_max_age = 2, 
      gen_is_variable = 0.,
      gen_full_load_heat_rate = ".",
      gen_is_baseload = 0L,
      gen_is_cogen = 0L,
      gen_energy_source	= "PHS",
      # gen_store_to_release_ratio = storage_units_max_hours,
      gen_store_to_release_ratio = '.',
      gen_storage_efficiency = 
        storage_units_efficiency_store * storage_units_efficiency_dispatch
       # gen_unit_size	gen_ccs_capture_efficiency	gen_ccs_energy_load
    )
  
  # gen_info_tech$gen_full_load_heat_rate[
  #   gen_info_tech$gen_is_variable == 1] <- "." 
  
  # gen_info_tech$gen_capacity_limit_mw[
  #   is.infinite(gen_info_tech$gen_capacity_limit_mw)] <- "."
  
  gen_info_phs$gen_tech |> unique()
  
  gen_info_csv <- 
    rbind(gen_info_csv, 
    select(gen_info_phs, GENERATION_PROJECT:gen_storage_efficiency), 
    use.names = T)
} else {
  gen_info_phs <- as.data.table(NULL)
}

```

### switch: gen_build_predetermined (PHS)

Pumped hydro storage, adding to `gen_build_predetermined` table.

```{r}
#| eval: !expr params$build_switch

if (nrow(gen_info_phs) > 0) {
  gen_build_predetermined_phs <- gen_info_phs |>
    filter(storage_units_p_nom > 0) |>
    mutate(
      build_year = p$base_year - 1,
      build_gen_predetermined = storage_units_p_nom
    ) |>
    select(GENERATION_PROJECT, build_year, build_gen_predetermined)
  
  # merge
  gen_build_predetermined <- 
    rbind(gen_build_predetermined,
          gen_build_predetermined_phs,
          use.names = T)
  stopifnot(!any(is.na((gen_build_predetermined))))
}
```

### switch: gen_build_costs (PHS)

```{r}
#| eval: !expr params$build_switch

if (nrow(gen_info_phs) > 0) {
  gen_build_costs_phs <- gen_info_phs |>
    mutate(
      build_year = NA,
      # build_year = if_else(extendable == 1,
      #                      p$base_year, # investable
      #                      p$base_year - 1 # not investable
      #                      ),
      gen_overnight_cost = 0,
      # gen_overnight_cost = ".",
      gen_fixed_om = 0L,
      gen_storage_energy_overnight_cost = storage_units_capital_cost
    ) |>
    select(GENERATION_PROJECT, build_year, gen_overnight_cost, gen_fixed_om,
           gen_storage_energy_overnight_cost)
  

  # costs of existing stock
  gen_build_costs_0 <- gen_build_costs_phs |>
    filter(GENERATION_PROJECT %in% gen_info_phs$GENERATION_PROJECT) |>
    mutate(
      build_year = p$base_year - 1
    )
  
  # investable 
  gen_build_costs_1 <- gen_build_costs_phs |>
    filter(
      GENERATION_PROJECT %in% 
        gen_info_phs$GENERATION_PROJECT[gen_info_phs$extendable == 1]) |>
    mutate(
      build_year = p$base_year
    )
  # merge
  gen_build_costs_phs <- rbind(gen_build_costs_0, gen_build_costs_1)
  rm(gen_build_costs_0, gen_build_costs_1)
  
  gen_build_costs_phs$GENERATION_PROJECT |> unique()
  stopifnot(!any(is.na((gen_build_costs_phs))))
  
  gen_build_costs <- 
    rbind(gen_build_costs,
          gen_build_costs_phs,
          use.names = T)
  print(gen_build_costs_phs)
}
```

## Hydro PP with a dam and exogenous inflow

This type of technology has both generating and storing capacity, but its charging process is defined by third factors - the inflow of the water into the dam. Therefore the discharge (electricity generation) can be optimized based on the given load and inflow profile, and the storing capacity.

```{r}
unique(nc$storage_units$storage_units_carrier)
hydro <- nc$storage_units |>
  filter(grepl("hydro", storage_units_carrier)) 
if (nrow(hydro) > 0) {
  hydro <- hydro |>
    mutate(
      # bus = str_extract(stores_bus, "[A-Z]+.[0-9]+"),
      bus = storage_units_bus,
      comm = carriers2comm(storage_units_carrier),
      stores = str_replace(storage_units_i, bus, ""),
      stores = str_trim(stores),
      stg = paste0("STG_", comm),
      sup = paste0("RES_", comm),
      tech = paste0("E", comm),
      weather = paste0("W", comm),
      extendable = 0L
      # extendable = if_else(storage_units_capital_cost > 0, 1, 0) #!!! assumed
    ) |>
    left_join(reg_names, by = "bus")
}

# hydro
```

### energyRt: (hydro PP with a dam, variant #1)

The most straightforward way to reproduce this type of process in energyRt is to model it with three classes: supply, storage, and technology.\
The supply class will introduce the hydro commodity in the model with a given availability profile. Parameter `@weather$wava.fx` will fix the supply to a given time series (see `inflow` table in *Hydro storage inflow*).

```{r}
#| eval: !expr params$build_energyRt

repo_hyddam <- newRepository("repo_hyddam", 
                             desc = "Hydro with dam - v1")
if (nrow(hydro) > 0) {
  
  #1. update commodity HYD
  HYD <- repo_comm@data$HYD
  repo_comm@data$HYD <- NULL # add to `repo_hyddam` repository
  if (is.null(HYD)) HYD <- newCommodity(name = "HYD", 
                                        desc = "Hydro energy")
  
  HYD <- update(HYD, 
                # limtype = "FX", # 
                timeframe = "HOUR"
                )
  # stopifnot(!any(unique(hydro$sup) %in% names(repo_sup@data))) ## check 
  
  for (s in unique(hydro$sup)) {
    x <- filter(hydro, sup == s)
    
    #2. supply with weather factors
    stopifnot(length(unique(x$weather)) == 1) # check for errors
    RES_HYD <- newSupply(
      name = s,
      commodity = x$comm[1],
      desc = paste0("Hydro resources"),
      availability = data.frame(
        region = x$region,
        ava.fx = 1 # pegged to the weather factor
      ),
      weather = list(
        weather = unique(x$weather),
        wava.fx = 1
      )
    )

    #3. generator (hydro turbine)
    EHYD <- newTechnology(
      name = "EHYD",
      desc = "Hydro turbine connected to a dam",
      input = list(comm = unique(x$comm)),
      output = list(comm = "ELC"),
      cap2act = 24*365,
      region = unique(x$region),
      ceff = list(
        region = x$region,
        comm = x$comm,
        cinp2use = x$storage_units_efficiency_dispatch
      ),
      capacity = list(
        region = x$region,
        stock = x$storage_units_p_nom 
      ),
      end = list(end = p$base_year - 1) #!!!ToDo: add checks
    )
    # draw(EHYD)

    #4. storage
    STG_HYD <- newStorage(
      name = "STG_HYD",
      desc = "Hydro dam",
      # cap2stg = 
      commodity = unique(x$comm),
      cap2stg = 6,
      region = unique(x$region),
      capacity = list(
        region = x$region,
        stock = x$storage_units_p_nom
      ),
      end = list(end = p$base_year - 1) #!!!ToDo add checks
    )
    
    repo_hyddam <- add(repo_hyddam, HYD, RES_HYD, EHYD, STG_HYD)
  }
  rm(HYD, RES_HYD, EHYD, STG_HYD)
  repo_hyddam@data |> names()
}
```

### switch: (hydro PP with a dam, variant #1)

This technology will be added on further steps

```{r, eval=params$build_switch}

```

## Chargers and dischargers ("links" in PyPSA)

```{r}
link2tech <- function(x, ignore.case = T) { #!!! ToDo: write a function
  # mapping of carriers to commodities
  x[grepl("H2.Electrolysis", x, ignore.case = ignore.case)] <- "ELC2H2"
  x[grepl("H2.Fuel.Cell", x, ignore.case = ignore.case)] <- "H2ELC"
  x[grepl("battery.charger", x, ignore.case = ignore.case)] <- "ELC2BTR"
  x[grepl("battery.discharger", x, ignore.case = ignore.case)] <- "BTR2ELC"
  x
}

links <- nc$links |>
  mutate(
    # bus = str_extract(links_i, "[A-Z]+.[0-9]+"),
    bus = rev_extract_bus(links_i),
    tech = link2tech(links_carrier),
    input = str_trim(str_replace(links_bus0, bus, "")),
    input = carriers2comm(input),
    input = if_else(input == "", "ELC", input),
    output = str_trim(str_replace(links_bus1, bus, "")),
    output = carriers2comm(output),
    output = if_else(output == "", "ELC", output)
    # tech_name = paste0("STG_", toupper(stores))
  ) |>
  left_join(reg_names, by = "bus") |>
  unique()

```

### energyRt: technologies ("links" in PyPSA)

```{r}
#| eval: !expr params$build_energyRt

repo_links <- newRepository("links")
if (nrow(links) > 0) {
  for (nm in unique(links$tech)) {
    x <- links |> filter(tech == nm)
    tech <- newTechnology(
      name = nm,
      desc = x$links_carrier[1],
      # desc = paste(x$links_i, collapse = ", "),
      input = list(comm = unique(x$input)),
      output = list(comm = unique(x$output)),
      cap2act = 24*365,
      region = unique(x$region),
      ceff = list(
        region = x$region,
        comm = x$input,
        cinp2use = x$links_efficiency
      ),
      olife = list(olife = 100), # doesn't matter for 1-year run and annualized inv
      fixom = list(
        region = x$region,
        fixom = x$links_capital_cost # annualized costs
      ),
      # invcost = list(
      #   region = x$region,
      #   invcost = x$links_capital_cost
      # ),
      end = list(
        region = x$region,
        end = if_else(x$links_p_nom_extendable == 0, 
                      p$base_year - 1, 
                      p$base_year + 1)
      )
    )
    repo_links <- add(repo_links, tech)
    rm(tech)
  }  
}
repo_links@data |> names()
# repo_links@data$ELC2H2 |> draw()
```

### switch: storage (accumulators + links)

```{r}
#| eval: !expr params$build_switch

# merge "links" and "storage" data tables, arranging by fuel type
acc_links <- links |>
  mutate(
    stores_carrier = str_extract(links_i, "H2|battery"),
    process = if_else(grepl("cell|discharger", links_carrier), "out", "inp")
  ) |>
  select(-(links_i:links_carrier), -(bus:output), -load_zones) |>
  unique() |>
  pivot_wider(
    names_from = c(process), 
    values_from = c(links_p_nom_extendable:links_capital_cost)) |>
  as.data.table() |>
  full_join(storage, by = c("region", "stores_carrier"))
stopifnot(all(!is.na((acc_links))))
```

### gen_info (accumulators + links)

Pumped hydro storage, adding to `gen_info` table.

```{r}
#| eval: !expr params$build_switch

if (nrow(acc_links) > 0) {
  gen_info_acc <- acc_links |>
    mutate(
      GENERATION_PROJECT = paste0(region, "_", stg),
      gen_tech = stg,
      gen_load_zone = region,
      gen_connect_cost_per_mw = 0.,
      gen_capacity_limit_mw = ".",
      gen_full_load_heat_rate = ".",
      gen_variable_om = 0.,
      gen_max_age = 2, # see also 
      gen_is_variable = 0.,
      gen_full_load_heat_rate = ".",
      gen_is_baseload = 0L,
      gen_is_cogen = 0L,
      gen_energy_source	= "Electricity", # == storage
      gen_store_to_release_ratio = ".",
      # gen_store_to_release_ratio = if_else(comm == "H2", 24 * 30, 6), # !!! assumption
      gen_storage_efficiency = 
        links_efficiency_inp  * links_efficiency_out
       # gen_unit_size	gen_ccs_capture_efficiency	gen_ccs_energy_load
    )
  
  gen_info_csv <-
    rbind(gen_info_csv, 
    select(gen_info_acc, GENERATION_PROJECT:gen_storage_efficiency), 
    use.names = T)
  stopifnot(all(!is.na((gen_info_csv))))
}

```

### switch: gen_build_costs (accumulators + links)

```{r}
#| eval: !expr params$build_switch

gen_build_costs_acc <- gen_info_acc |>
  mutate(
    build_year = p$base_year, # all in one
    # gen_overnight_cost = links_capital_cost_inp + links_capital_cost_out,
    gen_overnight_cost = 0L,
    gen_fixed_om = links_capital_cost_inp + links_capital_cost_out,
    gen_storage_energy_overnight_cost = stores_capital_cost * 
      gen_info_acc$gen_max_age[1] # annualized -> overnight (roughly)
      # 0L
      ## links_capital_cost_inp + links_capital_cost_out +
      # stores_capital_cost
    # * gen_store_to_release_ratio, # !!! assumption
  ) |>
  select(GENERATION_PROJECT, build_year, gen_overnight_cost, gen_fixed_om,
         gen_storage_energy_overnight_cost)
gen_build_costs_acc$GENERATION_PROJECT |> unique()

gen_build_costs <- gen_build_costs |>
  rbind(gen_build_costs_acc, use.names = T)
stopifnot(all(!is.na((gen_build_costs))))
```

```{r end_nc_import}
#| include: false
# close PyPSA-nc file
try(nc_close(nc), silent = T)
```

## (optional) Update PyPSA-nc

Update `p$files$pypsa_network` file with MERRA2 capacity factors and save in the same directory with `p$new_pypsa_file_ending`.

```{r}
#| eval: !expr params$update_pypsa_model

if (F) { # (optional) study nc-file structure
  # https://pjbartlein.github.io/REarthSysSci/netCDF.html
  # https://ropensci.org/blog/2019/11/05/tidync/
  library(tidync)
  src <- tidync(p$files$pypsa_network)
  print(src)
  src <- src |> activate("generators_t_p_max_pu")
  src |> activate("generators_i")
  src |> activate("generators_t_p_max_pu_i") |> hyper_tibble() |>
    as.data.table()
  src |> activate("D12") |> hyper_tibble() |>
  as.data.table()
}

if (p$update_pypsa_model) { # test writing to nc-file
  library(ncdf4)
  # new file name
  p$files$pypsa_network_merra <- 
    str_replace(p$files$pypsa_network, ".nc$", p$new_pypsa_file_ending)
  
  # copy nc-file (replacing previous copy with the same name)
  file.copy(from = p$files$pypsa_network, to = p$files$pypsa_network_merra,
            overwrite = TRUE)
  
  nc <- nc_open(p$files$pypsa_network_merra, write = TRUE)
  # nc_close(nc)
  # nc <- nc_open(file.path(p$dir$data_mod, 
  #                         "elec_s_10_ec_lcopt_Co2L-1H-merra.nc"), 
  #               write = T)
  attributes(nc)$names
  
  # capacity factors
  x <- ncvar_get(nc, "generators_t_p_max_pu")
  class(x); 
  dim(x); x[1:5, 1:5]; dimnames(x); 
  
  rownames(x) <- ncvar_get(nc, "generators_t_p_max_pu_i")
  # ncvar_get(nc, "snapshots")
  
  # update
  pypsa_update <- rbind(pypsa_update_onwind, pypsa_update_solar) |>
    pivot_wider(names_from = "snapshots", 
                values_from = "generators_t_p_max_pu")
  rnam <- pypsa_update$generators_t_p_max_pu_i
  cnam <- colnames(pypsa_update)[-1]
  stopifnot(all(as.integer(cnam) == ncvar_get(nc, "snapshots"))) # check
  pypsa_update <- pypsa_update |> select(-1) |> as.matrix()
  dim(pypsa_update)
  rownames(pypsa_update) <- rnam
  
  for (rn in rnam) {
    i <- which(rownames(x) == rn)
    x[i,] <- pypsa_update[rn,]
  }

  ncvar_put(nc, "generators_t_p_max_pu", x, verbose = T)
  nc_close(nc)
}
```

## switch: writing csv-files

### checks

Additional check to avoid switch error messages.

#### variable techs data

Check if variable capacity factors are available for all declared variable technologies.

```{r}
#| eval: !expr params$build_switch

# all variable technologies
var_gen <- gen_info_csv$GENERATION_PROJECT[gen_info_csv$gen_is_variable == 1]
stopifnot(anyDuplicated(var_gen) == 0) # no duplicates
var_gen <- unique(var_gen) |> sort()

# all capacity factors
var_cf <- unique(variable_capacity_factors$GENERATION_PROJECT) |> sort()

# check if all variable techs has been declared in `gen_info`
all_techs_declared <- all(var_cf %in% var_gen); all_techs_declared
stopifnot(all_techs_declared)

# check if capacity factors exist for all variable techs
all_cf_in_data <- all(var_gen %in% var_cf); all_cf_in_data
if (!all_cf_in_data) {
  # identify missing capacity factors
  ii <- !(var_gen %in% var_cf)
  var_gen[ii]
  if (params$force) {
    # try to resolve by dropping technologies with missing var-cap-factors
    ## 1. from `gen_info`
    jj <- gen_info_csv$GENERATION_PROJECT %in% var_gen[ii]
    if (any(jj)) {
      message("Missing technologies in `gen_info`: ",
              paste(gen_info_csv$GENERATION_PROJECT[jj], collapse = ", "))
      warning("dropping technologies without variable capacity factors: \n",
              paste(gen_info_csv$GENERATION_PROJECT[jj], collapse = ", "))
      gen_info_csv <- gen_info_csv[!jj,]
    }
    
    ## 2. from `gen_build_predetermined`
    jj <- gen_build_predetermined$GENERATION_PROJECT  %in% var_gen[ii]
    if (any(jj)) {
      message("Missing technologies in `gen_build_predetermined`: ",
              paste(gen_info_csv$GENERATION_PROJECT[jj], collapse = ", "))
      warning("dropping from `gen_build_predetermined`: \n",
              paste(gen_build_predetermined$GENERATION_PROJECT[jj], 
                    collapse = ", "))
      gen_build_predetermined <- gen_build_predetermined[!jj,]
    }

    ## 3. from `gen_build_costs`
    jj <- gen_build_costs$GENERATION_PROJECT %in% var_gen[ii]
    if (any(jj)) {
      tch_ii <- paste(gen_build_costs$GENERATION_PROJECT[jj], collapse = ", ")
      message("Missing info in `gen_build_costs` for:\n", tch_ii)
      warning("dropping from `gen_build_costs`:\n", tch_ii)
      gen_build_costs <- gen_build_costs[!jj,]
    }
    # gen_build_costs <- gen_build_costs[jj,]
  } else {
    # auto-correction is off - error exit
    stop("Missing variable_capacity_factors for: ",
         paste(var_gen[ii], collapse = " "))
  }
}
```

#### projects declared

```{r}
#| eval: !expr params$build_switch

# gen_build_predetermined
ii <- gen_build_predetermined$GENERATION_PROJECT %in%
  gen_info_csv$GENERATION_PROJECT
if (!all(ii)) {
  message("GENERATION_PROJECTs in gen_build_predetermined")
  print(gen_build_predetermined$GENERATION_PROJECT[!ii])
  message("... have not been declared in gen_info")
  # stop()
}

# gen_build_costs
ii <- gen_build_costs$GENERATION_PROJECT %in% gen_info_csv$GENERATION_PROJECT
if (!all(ii)) {
  message("GENERATION_PROJECTs in gen_build_predetermined")
  print(gen_build_costs$GENERATION_PROJECT[!ii])
  message("... have not been declared in gen_info")
  # stop()
}

```

### switch: gen_info.csv

```{r}
#| eval: !expr params$build_switch

stopifnot(all(!is.na((gen_info_csv))))
write.csv(gen_info_csv,
          file = file.path(p$scen$BASE$switch, "inputs/gen_info.csv"),
          row.names = FALSE, quote = FALSE)
```

### switch: gen_build_predetermined.csv

```{r}
#| eval: !expr params$build_switch

# add zeros to not-predetermined techs
all_projects <- gen_info_csv$GENERATION_PROJECT |> unique()
ii <- all_projects %in% unique(gen_build_predetermined$GENERATION_PROJECT)
gen_build_predetermined_0 <- data.table(
  GENERATION_PROJECT = all_projects[!ii],
  build_year = p$base_year,
  build_gen_predetermined = 0
)
gen_build_predetermined <- gen_build_predetermined |>
  # rbind(gen_build_predetermined_0, use.names = T) |>
  unique()
stopifnot(all(!is.na((gen_build_predetermined))))
write.csv(gen_build_predetermined,
          file = file.path(p$scen$BASE$switch, "inputs/gen_build_predetermined.csv"),
          row.names = FALSE, quote = FALSE)
```

### switch: gen_build_costs.csv

```{r}
#| eval: !expr params$build_switch
stopifnot(all(!is.na((gen_build_costs))))
write.csv(gen_build_costs,
         file = file.path(p$scen$BASE$switch, "inputs/gen_build_costs.csv"),
         row.names = FALSE, quote = FALSE)
```

### switch: financials.csv

```{r}
#| eval: !expr params$build_switch
financials <- tibble(
  base_financial_year	= periods$INVESTMENT_PERIOD,
  discount_rate	= p$discount, # 0.0, #5/100,
  interest_rate = p$discount #0.0, #7/100
)
stopifnot(all(!is.na((financials)))) 
write.csv(financials,
          file = file.path(p$scen$BASE$switch, "inputs/financials.csv"),
          row.names = FALSE, quote = FALSE)
financials
```

### switch: modules.txt

```{r}
#| eval: !expr params$build_switch
fl <- file.path(p$scen$BASE$switch, "inputs/modules.txt")
cat("# Core Modules", "\n", file = fl, append = FALSE)
cat("switch_model", "\n", file = fl, append = TRUE)
cat("switch_model.timescales", "\n", file = fl, append = TRUE)
cat("switch_model.financials", "\n", file = fl, append = TRUE)
cat("switch_model.balancing.load_zones", "\n", file = fl, append = TRUE)
cat("switch_model.energy_sources.properties", "\n", file = fl, append = TRUE)
cat("switch_model.generators.core.build", "\n", file = fl, append = TRUE)
cat("switch_model.generators.core.dispatch", "\n", file = fl, append = TRUE)
cat("switch_model.reporting", "\n", file = fl, append = TRUE)
cat("# Custom Modules", "\n", file = fl, append = TRUE)
cat("# switch_model.transmission.local_td", "\n", file = fl, append = TRUE)
cat("switch_model.generators.core.no_commit", "\n", file = fl, append = TRUE)
cat("# switch_model.energy_sources.fuel_costs.markets", "\n", file = fl, append = TRUE)
cat("switch_model.energy_sources.fuel_costs.simple", "\n", file = fl, append = TRUE)
cat("switch_model.transmission.transport.build", "\n", file = fl, append = TRUE)
cat("switch_model.transmission.transport.dispatch", "\n", file = fl, append = TRUE)
cat("switch_model.balancing.unserved_load", "\n", file = fl, append = TRUE)
cat("switch_model.generators.extensions.storage", "\n", file = fl, append = TRUE)

```

### switch: switch_inputs_version.txt

```{r}
#| eval: !expr params$build_switch
fl <- file.path(p$scen$BASE$switch, "inputs/switch_inputs_version.txt")
cat("2.0.7", "\n", file = fl, append = FALSE)

fl <- file.path(p$scen$BASE$switch, "options.txt")
cat("--verbose", "\n", file = fl, append = FALSE)
cat("--stream-solver", "\n", file = fl, append = TRUE)
cat("--inputs-dir inputs", "\n", file = fl, append = TRUE)
cat("--no-save-solution", "\n", file = fl, append = TRUE)
cat("--solver cplex", "\n", file = fl, append = TRUE)
cat("--solver-options-string 'lpmethod=4 solutiontype=2'", "\n", file = fl,
    append = TRUE)
rm(fl)
```

### switch: solve

```{r}
#| eval: !expr params$build_switch

# "conda info --envs"
fl <- file.path(p$scen$BASE$switch, "run_switch.cmd")
# cat("cd ", normalizePath(dirname(fl)),"\n", file = fl, append = FALSE)
cat("call activate switch", "\n", file = fl, append = FALSE)
cat("switch solve --verbose --solver cplex --full-traceback --stream-solver",
    "\n", file = fl, append = TRUE)
cat("call conda deactivate", "\n", file = fl, append = TRUE)
cat("cmd /k", "\n", file = fl, append = TRUE)

print("switch: solve")
# plan(multisession(workers = 2))
# switch_eval <- future({
# system(fl, wait = F, invisible = F)
# print("switch solve started")
# })
    
# alternatively use the command in conda/python to run switch model:
"switch solve --verbose --solver cplex --full-traceback --stream-solver"
```

## energyRt: model

### (optional) cost of unserved load

Capacity expansion models (energy systems in general) take load (final demand) as a given, exogenous level of electricity (demand-commodity) to deliver. Optimization of supply-side, capacity of generation and balancing technologies (and flexible demand structure when available) to serve the given level of load at any costs in every hour and every region. In the case when there is not enough resources to serve the load, the solution will be infeasible, the model will not converge to optimal solution.\
To avoid this type of infeasibility and also study narrow places in the modeled resources, potential bags in the data, it is recommended to supply a cost of not delivering (unserving) the load. Hours and regions where the load is unserved indicate bottlenecks in the model, and subject for further investigation of the data and constraints.\
`Switch` has an "switch_model.balancing.unserved_load" module that sets default 500 USD/MWh price on unserved load. (see above)\
In `energyRt` one can put a dummy-technology or import of electricity with a certain cost that will be used if marginal production costs of electricity in particular hour and region will exceed the price. Below is an example of an unserved load modeled as an import.

```{r, eval=FALSE}
#| eval: !expr params$build_energyRt

EIMP5 <- newImport(
  name = "EIMP",
  desc = "Unserved load 5 USD per kWh",
  commodity = "ELC",
  imp = list(price = 500) 
)

EIMP10 <- newImport(
  name = "EIMP",
  desc = "Unserved load 10 USD per kWh",
  commodity = "ELC",
  imp = list(price = 1000) 
)

EIMP100 <- newImport(
  name = "EIMP",
  desc = "Unserved load 100 USD per kWh",
  commodity = "ELC",
  imp = list(price = 10000)
)

```

### energyRt: assemble model

```{r, eval=FALSE}
#| eval: !expr params$build_energyRt

repo <- newRepository("base") |>
  add(
    repo_comm, # main commodities
    repo_sup,  # supply/resources
    repo_gen,  # generators
    repo_network, # inter-regional network
    repo_stg, # storage (accumulators)
    repo_phs, # pumped hydro
    repo_hyddam, # hydro with a dam
    repo_links, # chargers and dischargers (links) for storage
    repo_inflow, # hydro inflow
    repo_ror_cf, # run of river capacity factors
    repo_onwind_cf, # onshore wind capacity factors
    repo_solar_cf, # solar capacity factors
    EIMP10, # optional price of unserved load
    DEMELC # load curve
  )

# Optional: all commodities set to "HOUR" timeframe"
if (F) {
  for (ob in repo@data) {
    if (.hasSlot(ob, "timeframe") && length(ob@timeframe) > 0 &&
        grepl("ANNUAL|YDAY", ob@timeframe)) {
      ob <- update(ob, timeframe = "HOUR")
      repo <- add(repo, ob, overwrite = T)
    }
  }
}

timetable_d365_h24 <- make_timetable(
  struct = list(
    ANNUAL = "ANNUAL",
    YDAY = str_extract(timedim$slice, "d[0-9]+") |> unique(),
    HOUR = str_extract(timedim$slice, "h[0-9]+") |> unique()
  )
)

calendar_d365_h24 <- newCalendar(timetable_d365_h24, name = "calendar_d365_h24")


mod <- newModel(
  name = p$scen$BASE$name,
  desc = "Reproduced PyPSA model with MERRA2 cf in energyRt",
  region = reg_names$region,
  discount = p$discount,
  calendar = calendar_d365_h24,
  repository = repo
)

mod <- setHorizon(mod, p$base_year, 1)
mod@config@horizon

save(mod, file = file.path(p$scen$BASE$energyRt, "energyRt_model.RData"))
```

### energyRt: solve

```{r, eval=FALSE}
#| eval: false
#| include: false
# eval: !expr params$solve_energyRt
# show_progress_bar()
set_progress_bar("bw")
scen_pypsa_base <- interpolate(
  mod, 
  name = p$scen$BASE$name,
  path = p$scen$BASE$energyRt
)
scen_pypsa_base@path

```
