---
title: "Wind and Solar capacity factors based on MERRA2 data"
date: "`r format(Sys.time(), tz = 'UTC', format = '%Y-%m-%d %H:%M:%S %Z')`"
# output: html_document
output: 
  pdf_document: 
    toc: yes
    toc_depth: 5
    fig_caption: yes
  html_document: 
    - rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{merra}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
params: 
  # wind technology
  # select wind locations
  # cluster wind location 
  # maximum loss of s.d. by clustering (ignored if win_cl_max = 1)
  # solar technology
  # solar locations - not implemented
  # cluster solar locations
  # iso3c: BGD # country code 
  iso3c: VNM
  # yaml_path: "config/mod_adm" #  file: "config_{iso3c}.yml
  # yaml_path: "config/mod_toy" #
  yaml_path: "config/mod_custom" #
  weather_year: 2013 # weather year
  reimport_all: false # if true, saved merra and pre-processed data won't be used
  win_hub_height: 100 # meters, hub height, options:  50, 100, 150 (see info) 
  win_onshore_max_MW_km2: 4 # assumption (see info), maximum MW per km2
  win_offshore_max_MW_km2: 4 # assumption (see info), maximum MW per km2
  win_gwa_regroup: false # if true, redo GWA groupping
  win_gwa_cf_min: 0.2 # lowest capacity factor in Global Wind Atlas, class 2
  win_merra_cf_min: 0.1 # lowest capacity factor according to MERRA2
  win_recluster: false # ignore/overwrite previously saved clusters
  # win_cl_max: null # maximum number of clusters of wind-data per region
  win_cl_tol: [0.02, .05, 0.1, 0.2, 1.0] 
  sol_pv_type: tl # tracker options: fh, fl, tl, td (see info)
  sol_onshore_max_MW_km2: 80 # assumption (see info), maximum MW per km2
  sol_offshore_max_MW_km2: 0 # assumption (see info), maximum MW per km2
  sol_recluster: false # ignore/overwrite previously saved clusters
  # sol_cl_max: null # maximum number clusters of solar-data per region
  sol_cl_tol: [0.01, 0.02, .05, 0.1, 1.0]  # maximum loss of s.d. 
  echo: false # show R-script in reports
  print_parameters: true # print YAML parameters
  file_prefix: "merra_"
  file_suffix: "_{param$weather_year}"
editor_options: 
  markdown: 
    wrap: 72
  chunk_output_type: console
---

```{r knitr_setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(
  echo = params$echo, 
  knitr.graphics.auto_pdf = T, 
  warning = F,
  message = F,
  fig.width = 6, 
  fig.height = 6
  )
```

 

::: {.center data-latex=""}
**`r countrycode::countrycode(params$iso3c, "iso3c", "country.name")`**\
**`r paste0("model settings: ", params$yaml_path, glue("/config_{params$iso3c}.yml"))`**
:::

## Info

This file has been used to create capacity factors from MERRA2 dataset for each model-region and cluster them based on temporal correlation of neighbor cells with pre-set level of information loss due to the aggregation. (see `merra2ools` R-package for details.)
[the script below is intened to be run chunk-by-chunk, the "polishing" and migration of the script to .qmd file is not finished]

### Output

-   **gwa_iec2_sf.RData** -- saved `gwa_iec2_sf`, an `sf` object (the
    model map) of grouped by level wind capacity factors (GWA, class
    *IEC2*). Columns:
    -   `int` and `eq` indicate the capacity factor group,
    -   `area` - estimated area of the group (km\^2),
    -   ...
-   **locid_gwa_sf.RData** -- saved `locid_gwa_sf` object, which is a
    merged GWA levels (`gwa_iec2_sf`) with MERRA2 summary data:
    -   `locid` location identifiers (from `merra2ools` package),
    -   `wcf50m`, `wcf100m`, `wcf150m` -- 40-years averages of estimated
        capacity factors for hub hights 50, 100, and 150 meters
        respectively.
-   **locid_wcf{win_hub_height}m_cl.RData** -- file in the data
    directory (`p$dir$data`) with

\newpage

### Names convention

Parameters, objects, files, and paths in the document and project use
the same naming convention:

-   **merra\_** source of the original data is MERRA2
-   **locid\_** MERRA2-related data with location identifier (`locid` in
    `merra2ools` package)
-   **win\_** wind data and related files
-   **sol\_** solar data and related files
-   **gwa** used data from Global Wind Atlas (GWA) - **\_sf** object of
    *simple feature* (`sf`) class, with spatial data assigned to every
    row of the table.
-   **\_min** or **\_max** lower or upper bound on a parameter
-   **fname** name of a file
-   **mask** an identifier of a certain (clustering) methodology in a
    file name.
-   **region** model region with offshore areas (if available) assigned
    to the closest land region.
-   **reg_off** names of regions where every offshore has its own
    identifier, with **\_off** ending. - **cl** clustered data
-   **TOL{:NUMBER:}** in clustered data, an indicator of tolerance level
    of maximum standard deviation loss from clusterization. The
    tolerance is set in the YAML header of the document, from 0 to 1,
    the tolerance level mark is reported in percent. Examples:
    -   `_TOL01_` is up to 1% loss of s.d.;
    -   `_TOL05_`, `_TOL10_`, `_TOL20_` -- 5%, 10%, and 20% tolerance
        respectively;
    -   `_TOL00_` is 0% loss of s.d., meaning the number of clusters is
        equal to the number of observations;
    -   `_TOL99_` is 99% \~= 100% loss tolerance, meaning 1 cluster per
        region;

------------------------------------------------------------------------

## Setup
Directories:  
```{r setup, warning=FALSE, message=FALSE}
## Load libraries
library(tidyverse)
library(revaluation)
# devtools::install_github("energyRt/merra2ools")
library(merra2ools)
# library(SP)
library(sf)
library(data.table)
library(fst)
library(glue)
library(terra)
library(globalwindatlas)
library(TSdist)
source("functions.R")

# read configuration file for the country
p <- rev_read_yaml(path = params$yaml_path, iso3c = params$iso3c)
# import/merge YAML parameters
for (i in 1:length(params)) {p[[names(params)[i]]] <- params[[i]]}

# Read cluster-parameters
## to avoid multiple instances of the same cluster parameters
# if (!is.null(p$win_cl_max) && p$win_cl_max == 1) {p$win_cl_tol <- 1} 
# if (!is.null(p$sol_cl_max) && p$sol_cl_max == 1) {p$sol_cl_tol <- 1} 

## cluster-version identifier in file-names of capacity factors 
# win_cl_mask <- 
#   glue(
#     if_else(is.null(p$win_cl_max), "", c("CL", params$win_cl_max)), 
#     if_else(!is.null(p$sol_cl_max) & !is.null(p$win_cl_tol), "_", ""),
#     "TOL", prettyNum(p$win_cl_tol) # loss/tolerance
#   ) %>% 
#   str_replace_all("\\.", "")

# sol_cl_mask <- 
#   glue(
#     if_else(is.null(p$sol_cl_max), "", c("CL", params$sol_cl_max)), 
#     if_else(!is.null(p$sol_cl_max) & !is.null(p$sol_cl_tol), "_", ""),
#     "TOL", prettyNum(p$sol_cl_tol) # loss/tolerance
#   ) %>% 
#   str_replace_all("\\.", "")

# Decode solar & wind type of technologies in MERRA-2 data
p$win_cf <- glue("wcf{params$win_hub_height}m")
p$sol_cf <- glue("scf_{params$sol_pv_type}")

# load maps
(load(p$files$gis))
gis_sf <- gis$mod_map$sf 
p$nreg <- gis_sf$region %>% unique() %>% length()
p$nreg_off <- gis_sf$reg_off %>% unique() %>% length()

wyear <- p$weather_year # short version 

if (is.null( p$fig_map_width) | is.null( p$fig_map_height)) {
  # propose parameters for map-figures
  b <- st_bbox(gis$mod_map$sf) 
  fig_rat <-  (b[4] - b[2]) / (b[3] - b[1])
  if (fig_rat < 1) { # horizontal
    p$fig_map_width <- 6
    p$fig_map_height <- round(4 * fig_rat * p$fig_map_width) / 4
  } else { # vertical
    p$fig_map_height <- 6
    p$fig_map_width <- min(6, max(4, round(p$fig_map_height / fig_rat * 4) / 4))
  }
}

ggsave_map <- function(filename = default_name(plot), 
                       path = p$dir$fig,
                       # height = p$fig_map_height,
                       # width = p$fig_map_width,
                       # scale = 1.25,
                       ...) {
  ggsave(filename = filename, height = p$fig_map_height, 
         width = p$fig_map_width, 
         scale = 1.25, path = path, ...)
}

if (F) { # test figure ratio
  
  a <- ggplot(gis_sf) +
    geom_sf(aes(fill = region), show.legend = T) +
    scale_fill_viridis_d(option = "H") +
    # theme_void()
    theme_bw()
  
  ggsave("test_map_ration.png", plot = a, 
         # height = 6, width = 5, # THA, BGD, CHL
         # height = 5, width = 5, # COL
         # height = 5, width = 6, # IND, CHN
         # height = 3.8, width = 6.5, # KAZ
         # height = 3., width = 6., # MYS
         # height = 6, width = 4., # VNM
         height = p$fig_map_height, width = p$fig_map_width,
         scale = 1.25, path = p$dir$tmp)
  fs::file_show(path = file.path(p$dir$tmp, "test_map_ratio.png"))
}

```

\newpage

## Locations

```{r locid_all}
# identify location identifiers in merra2ools dataset for the model map
locid_all <- get_locid(gis_sf)
p$n_locid <- length(locid_all)

# get 40-years average capacity factors (& rename "af" to "cf" for consistency)
mean_cf <- merra2ools::mean_cf
nms <- names(mean_cf) %>% str_replace("af", "cf"); names(mean_cf) <- nms; rm(nms)

# create grid of selected locations
locid_grid_sf <- get_merra2_grid("poly", locid = locid_all) %>%
  select(locid) %>%
  left_join(mean_cf) %>%
  st_make_valid()

ggplot() +
  geom_sf(aes(fill = region), alpha = .5, data = gis_sf) +
  geom_sf(fill = NA, data = locid_grid_sf, color = "red") +
  labs(title = glue("MERRA2 grid, ", p$n_locid, " locations")) +
  rev_theme_map()
ggsave_map("merra2_grid.png")
```

```{r fig_locid_grid_sf_sol}
ggplot() + 
  geom_sf(aes(fill = .data[[p$sol_cf]]), data = locid_grid_sf) +
  scale_fill_viridis_c(option = "A") +
  geom_sf(data = gis_sf, fill = NA, color = "white") +
  labs(title = "Solar capacity factors, 40-years average") +
  rev_theme_map()
ggsave_map(glue("locid_grid_sf-{p$sol_cf}.png"))
```

```{r fig_locid_grid_sf_win}
ggplot() + 
  geom_sf(aes(fill = .data[[p$win_cf]]), data = locid_grid_sf) +
  scale_fill_viridis_c(option = "D") +
  geom_sf(data = gis_sf, fill = NA, color = "white") +
  labs(title = "Wind capacity factors, 40-years average") +
  rev_theme_map()
ggsave_map(glue("locid_grid_sf-{p$win_cf}.png"))
```

\newpage

## MERRA2 subset

```{r merra_subset, eval=TRUE}
fname <- file.path(p$dir$merra2, glue("merra_raw_{wyear}.fst"))
if (!file.exists(fname) | p$reimport_all) {
  print(glue("Importing data from MERRA2 database, ", p$n_locid,
          " locations, ", p$weather_year, " year"))
  merra <- get_merra2_subset(locid = locid_all, 
                             from = fDate(wyear, 01, 01, 0), 
                             to = fDate(wyear, 12, 31, 23), 
                             tz = "UTC")
  print(glue("Saving 'raw' data: ", fname))
  write_fst(merra, fname, compress = 100)
  cat("File size:", {file.size(fname) %>% gdata::humanReadable()})
} else {
  print(glue("Reading MERRA2 subset: ", fname))
  merra <- read_fst(fname, as.data.table = T)
}
```

## Wind

Process MERRA2 data (subset) to estimate hourly wind capacity factors
and group locations.

### Capacity factors

Estimating hourly capacity factors for selected year & locations. The
algorithm is using generic wind power curve to estimate potential output
based on the wind speed at a give height. Wind speeds at heights other
than 50 meters will be extrapolated using wind gradient function and
Hellman constant, estimated on wind speed at 10 and 50 meters height,
available in MERRA2 dataset.

```{r merra_wind, eval=TRUE}
merra_wind <- merra %>% 
  fWindCF(50, return_name = "wcf50m") %>% # wind
  fWindCF(100, return_name = "wcf100m") %>%
  fWindCF(150, return_name = "wcf150m") %>%
  select("UTC", "locid", starts_with("wcf"))
merra_wind
```

### Select locations

Global Wind Atlas offers high-resolution maps of wind potential. This
section filters locations with high wind capacity factors.

```{r gwa}
# gwa_set_dir(file.path(p$dir$data, "gwa"))
# gwa_get_dir()
# 
# gwa_tif <- gwa_get_wind_cf(p$iso3c, IEC = 2) # wind-class #2
# 
# ## plot 
# plot(rast(gwa_tif), main = "GWA capacity factors, IEC class 2")
```

### Group locations

```{r group_gwa}
fname <- file.path(p$dir$data_mod, "gwa_iec2_sf.RData")
if (!file.exists(fname) | p$win_gwa_regroup) {
  # set `glogalwindatlas` directory
  gwa_set_dir(file.path(p$dir$data, "gwa"))
  gwa_get_dir()
  
  # get (saved or download) tiff-file
  gwa_tif <- gwa_get_wind_cf(p$iso3c, IEC = 2) # wind-class #2
  
  ## save plot 
  png(glue("{p$dir$fig}/gwa_{p$iso3c}.png"), units = "in", res = 300,
      width = p$fig_map_width, height = p$fig_map_height)
  plot(rast(gwa_tif), main = "GWA capacity factors, IEC class 2")
  dev.off()
  
  # group locations by capacity factor
  gwa_iec2_sf <- gwa_group_locations(gwa_tif, 
                                     gis_sf, 
                                     aggregate_tif = 0, 
                                     snap_to_grid = 1e-05,
                                     drop_crumps = 100,
                                     simplify = 0.001,
                                     buffer = 0,
                                     int = seq(0, 1, by = .05), 
                                     verbose = T)
  save(gwa_iec2_sf, file = fname); rm(fname)
} else {
  load(fname); rm(fname)
}

# plot grouping results
ggplot(gwa_iec2_sf) +
  geom_sf(aes(fill = eq), color = NA) +
  scale_fill_viridis_d(option = "H", direction = 1, name = "CF") +
  labs(title = "Grouped wind capacity factors", 
       subtitle = "(Global Wind Atlas)") +
  rev_theme_map()
ggsave_map("gwa_iec2_sf.png")
```

### Filter locations

```{r filter_locid}
# intersect selected GWA locations with merra-grid 
# and filter out locations with low wind potential according to GWA data
locid_gwa_sf <- locid_grid_sf %>%
  st_make_valid() %>%
  st_intersection(
    filter(gwa_iec2_sf, int >= p$win_gwa_cf_min),
    dimensions = c("polygon")
  ) %>%
  st_make_valid() %>%
  mutate(
    area = units::set_units(st_area(geometry), "km^2"),
    win_MW_max = as.numeric(round(if_else(offshore, 4 * area, 4 * area), 0)),
    .before = "geometry"
  )

ggplot(locid_gwa_sf) +
  geom_sf(color = alpha("black", .25), fill = "lightgrey", data = gis_sf) +
  geom_sf(aes(fill = .data[[p$win_cf]]), color = NA) +
  scale_fill_viridis_c(option = "H", direction = 1, name = "waf100m") +
  labs(title = paste0(p$country, ", wind locations with cf >= ", 
                      p$win_gwa_cf_min, " (GWA)")) +
  rev_theme_map()
ggsave_map("locid_gwa_sf.png", path = p$dir$fig_mod)

# filter out location with low wind potential according to MERRA2 data
ii <- locid_gwa_sf[[p$win_cf]] >= p$win_merra_cf_min; 
print(glue("Excluded locations: {sum(!ii)} (out of {length(ii)})"))
# summary(ii)
locid_win_sf <- locid_gwa_sf %>% filter(ii); rm(ii)

ggplot(locid_win_sf) +
  geom_sf(color = alpha("black", .25), fill = "lightgrey", data = gis_sf) +
  geom_sf(aes(fill = .data[[p$win_cf]]), color = NA) +
  scale_fill_viridis_c(option = "H", direction = 1, name = "CF 100m\n(MERRA)") +
  labs(title = paste0(p$country, ", filtered wind locations"),
       subtitle = glue(
          "GWA cf >= {p$win_gwa_cf_min},", 
          " MERRA cf >= {p$win_merra_cf_min}"
          )) +
  rev_theme_map()
ggsave_map("locid_win_sf.png", path = p$dir$fig_mod)
```

### Add locations

(optional) add MERRA2 locations, beyond GWA map/data. [not implemented
yet]

```{r add_win_locations}

```

### Cluster locations

Use correlation-based algorithm for time series to cluster and group
MERRA2-locations.

```{r cluster_win, eval=T}
fname <- file.path(p$dir$data_mod, glue("locid_{p$win_cf}_cl.RData"))
if (!file.exists(fname) | p$win_recluster) {
  # number of clusters to consider
  if (p$n_locid / p$nreg > 50) { 
    # reducing computation time of clusters by reducing the number of options
    p$win_k_seq <- c(1:10, 12, 15, 20, 25, 35, 50, 70, 100, 150, 
                     seq(200, 1000, by = 100))
  } else {
    p$win_k_seq <- NULL # use default settings
  }
  # cluster locations
  locid_win_cl <- cluster_locid(merra_wind, 
                  varname = p$win_cf, 
                  locid_info = locid_win_sf, 
                  group = "reg_off", 
                  weight = "win_MW_max",
                  # k = 1:p$win_cl_max,
                  # k = 1:10,
                  k = p$win_k_seq,
                  # max_loss = p$win_cl_tol,
                  max_loss = .01,
                  verbose = T)
  print(glue("Saving clustering results: ", fname))
  save(locid_win_cl, file = fname); rm(fname)
} else {
  print(glue("Loading: ", fname))
  load(fname); rm(fname)
}

```

### Aggregate

**merra_win_cl** - hourly wind capacity factors, aggregated by cluster &
region.\
**locid_win_cl_sf** - spatial data.frame with location identifiers (locid) and clusters for each region of the model \
**win_GW_max_sf** - spatial data.frame with clusters by region and 
estimated maximum installed capacity, based on surface area (see
YAML-header parameters).

```{r aggregate_win, fig.height=8., warning=FALSE}
for (tol in p$win_cl_tol) {
  print(paste0("Aggregating MERRA2 location using clusters with tol = ", tol))
  try({
    # (suppressWarnings({
  # tol <- p$win_cl_tol[1]
  # stop()
  # create file identifier for clustered data
  win_cl_mask <- rev_merra_cl_filemask(tol)
  
  # Select clusters based on tolerance parameter (see yaml header)
  locid_win_cl_k <- locid_win_cl %>% 
      filter(sd_loss <= tol) %>% 
      group_by(reg_off) %>% 
      mutate(k_min = (k == min(k))) %>% ungroup() %>% 
      filter(k_min) %>% select(-k_min) %>%
      as.data.table()
  k_max <- max(locid_win_cl_k$k) # for figures
  nreg_data <- length(unique(locid_win_cl_k$reg_off)) # for figures
  # ToDo: split figures by ~20 groups for more than 20 regions
  rr <- unique(locid_win_cl_k$reg_off)
  if (length(rr) > 20) {
    rr <- rr[1:20]
    print(paste0("Showing results for first 20 regions (out of ", 
          length(unique(locid_win_cl_k$reg_off)), ")"))
  }
  # Cluster-loss figure
  locid_win_cl_kk <- locid_win_cl %>%
    group_by(reg_off, k) %>%
    summarise(sd_loss = max(sd_loss), N = max(N), .groups = "drop")
  
  locid_win_cl_k_i <- locid_win_cl_k %>%
    group_by(reg_off, k) %>%
    summarise(sd_loss = max(sd_loss), N = max(N), .groups = "drop")
  
  a <- ggplot(filter(locid_win_cl_kk, reg_off %in% rr)) +
    geom_line(aes(k, sd_loss), color = "dodgerblue", linewidth = 1.5) +
    geom_point(aes(k, sd_loss), color = "red", 
               data = filter(locid_win_cl_k_i, reg_off %in% rr)) +
    # geom_point(aes(k, sd_loss), color = "red", data = locid_win_cl_k_i) +
    scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
    scale_x_continuous(breaks = rev_integer_breaks(5)) +
    labs(x = "Number of clusters (k)", 
         y = "loss, % of s.d.",
         title = "Aggregation lossess by number of clusters of wind sites") +
    facet_wrap(~paste0(reg_off, " (", N," locations)"), scales = "free",
               ncol = ) +
               # ncol = max(1, round(p$nreg_off / 5))) +
    rev_theme_map()
  
  if (tol < 1) {
    a <- a + 
      geom_hline(aes(yintercept = tol), col = "red", linetype = 2) +
      labs(subtitle = glue("Targeted tolerance ", 100 * tol, "% ",
                          "and minimal sufficient number of clusters",
                          " (red line and dot)"))
  } else {
    a <- a + 
      labs(subtitle = 
             glue("One cluster (sd_loss tolerance = ", 100 * tol, "%)"))
  }
  print(a)
  ggsave(glue("win_cl_sd_loss_{win_cl_mask}.png"), a, path = p$dir$fig_mod, 
         width = 6, height = 8.5, scale = 1.25)
  
  # merging clusters of targeted tolerance with locid map
  fname <- file.path(p$dir$data_mod, glue("locid_win_cl_sf_{win_cl_mask}.RData"))
  
  # suppressMessages({
  locid_win_cl_sf <- locid_win_sf %>%
    st_make_valid() %>%
    left_join(
      select(locid_win_cl_k, any_of(c("reg_off", "region", "locid", "cluster")))
    ) %>%
    filter(!is.na(cluster)) %>%
    mutate(cluster = as.integer(cluster))
  # })
  print(glue("Writing ", fname))
  save(locid_win_cl_sf, file = fname); rm(fname)
  
  a <- ggplot(locid_win_cl_sf) +
    geom_sf(fill = "lightgrey", data = gis_sf)
  if (k_max > 10) {
    a <- a + geom_sf(aes(fill = as.integer(cluster)), color = NA) +
    scale_fill_viridis_c(option = "D", direction = 1, name = "Cluster")  
  } else {
    a <- a + geom_sf(aes(fill = as.factor(cluster)), color = NA) +
    scale_fill_viridis_d(option = "D", direction = 1, name = "Cluster")
  }
  a <- a +
    geom_sf(color = alpha("black", 1), fill = NA, data = gis_sf) +
    labs(title = paste0("Clustered wind sites by region",
                        ", sd_loss <= ", tol * 100, "%")) +
    rev_theme_map()
  # suppressMessages({
  print(a); rm(k_max)
  # ggsave_map(glue("locid_win_cl_sf_{win_cl_mask}.png"), a)
  ggsave(glue("locid_win_cl_sf_{win_cl_mask}.png"), a, 
       path = p$dir$fig_mod, scale = 1.25, width = p$fig_map_width, 
       height = p$fig_map_height)  
  # })
  
# save clustered hourly capacity factors
  # suppressMessages({
  merra_win_cl <- merra_wind %>%
    filter(locid %in% unique(locid_win_cl$locid)) %>% # drop unused locations
    mutate(slice = timeslices::dtm2tsl(UTC)) %>% # create time-slice IDs
    select(locid, slice, wcf50m, wcf100m, wcf150m) %>% # drop unused columns
    right_join( # merge with gis-data
      select(st_drop_geometry(locid_win_cl_sf), 
             locid, region, reg_off, offshore, cluster, win_MW_max), 
      relationship = "many-to-many"
    ) %>%
    group_by(slice, region, reg_off, offshore, cluster) %>%
    summarise(
      wcf50m = round(weighted.mean(wcf50m, win_MW_max, na.rm = T), 3),
      wcf100m = round(weighted.mean(wcf100m, win_MW_max, na.rm = T), 3),
      wcf150m = round(weighted.mean(wcf150m, win_MW_max, na.rm = T), 3),
      .groups = "drop"
    ) %>%
    mutate(year = p$weather_year, .before = 1) %>%
    select(-region, -offshore) %>%
    as.data.table()
  # })
  merra_win_cl
  
  fname <- file.path(p$dir$data_mod, 
                     glue("merra_win_{win_cl_mask}_{wyear}.fst"))
  print(glue("Writing ", fname))
  write_fst(merra_win_cl, fname, compress = 100); rm(fname)
  
  win_GW_max_sf <- locid_win_cl_sf %>% 
    filter(win_MW_max > 0) %>%
    st_make_valid() %>%
    # select(region, reg_off, offshore, cluster, win_MW_max) %>% 
    group_by(region, reg_off, offshore, cluster) %>% 
    reframe(
      wcf50m = weighted.mean(wcf50m, win_MW_max, na.rm = T),
      wcf100m = weighted.mean(wcf100m, win_MW_max, na.rm = T),
      wcf150m = weighted.mean(wcf150m, win_MW_max, na.rm = T),
      area = sum(area, na.rm = T),
      win_GW_max = sum(win_MW_max, na.rm = T) / 1e3,
      geometry = rev_union_polygons(geometry)
    ) %>% 
    st_as_sf() %>% st_make_valid()
  fname <- file.path(p$dir$data_mod, glue("win_GW_max_{win_cl_mask}.RData"))
  print(glue("Writing ", fname))
  save(win_GW_max_sf, file = fname); rm(fname)
  
  a <- ggplot(win_GW_max_sf) +
    geom_sf(fill = "lightgrey", data = gis_sf) +
    geom_sf(aes(fill = .data[[p$win_cf]]), color = alpha("black", .5),) +
    scale_fill_viridis_c(option = "H", direction = 1) +
    labs(title = "Average wind capacity factors by cluster and region",
         subtitle = glue("sd_loss <= ", tol * 100, 
                         "%, data: MERRA2, year: {wyear}")) +
    rev_theme_map()
  # suppressMessages({
  print(a)
  ggsave(glue("win_GW_max_sf_{win_cl_mask}_{wyear}.png"), a, 
         path = p$dir$fig_mod, width = p$fig_map_width, height = p$fig_map_height, 
         scale = 1.25)
  })
  # })
}
```

\newpage

## Solar

Process MERRA2 subset to estimate hourly solar capacity factors, group
and aggregate locations.

### Capacity factors

```{r merra_solar, eval=TRUE}
merra_solar <- merra %>% 
  fPOA(array.type = c("fh", "fl", 
                      # "th", "tv", # rarely used & similar to other types
                      "tl", "td")) %>% # solar (Plain of Array irradiance)
  mutate(
    # simplified version of capacity factors, 
    # assuming pick of output when POA >= 1000 Watt/m^2
    scf_fh = round(POA.fh / 1e3, 3),
    scf_fl = round(POA.fl / 1e3, 3),
    # scf_th = round(POA.th / 1e3, 3),
    # scf_tv = round(POA.tv / 1e3, 3),
    scf_tl = round(POA.tl / 1e3, 3),
    scf_td = round(POA.td / 1e3, 3)
  ) %>%
  mutate(
    # curtail cf > 1 (when POA > 1kW/m^2)
    scf_fh = if_else(scf_fh > 1, 1, scf_fh),
    scf_fl = if_else(scf_fl > 1, 1, scf_fl),
    # scf_th = if_else(scf_th > 1, 1, scf_th),
    # scf_tv = if_else(scf_tv > 1, 1, scf_tv),
    scf_tl = if_else(scf_tl > 1, 1, scf_tl),
    scf_td = if_else(scf_td > 1, 1, scf_td)
  ) %>%
  select(locid, UTC, starts_with("scf_"))
# summary(merra_solar)
```

### Locations

```{r locid_sol_sf, warning=FALSE}
locid_sol_sf <- locid_grid_sf %>%
  st_make_valid() %>%
  st_intersection(st_make_valid(gis_sf)) %>%
  st_make_valid() %>%
  mutate(
    area = units::set_units(st_area(geometry), "km^2"),
    sol_MW_max = as.numeric(round(if_else(
      offshore, 
      p$sol_offshore_max_MW_km2 * as.numeric(area),
      p$sol_onshore_max_MW_km2 * as.numeric(area)), 0)),
    .before = "geometry"
  )
# plot(locid_sol_sf["locid"])
```

### Filter locations

[reserved - not implemented]

```{r , eval=TRUE}

```

### Cluster locations

```{r sol_cluster, eval=TRUE}
fname <- file.path(p$dir$data_mod, glue("locid_{p$sol_cf}_cl.RData"))
if (!file.exists(fname) | p$sol_recluster) {
  # number of clusters to consider
  if (p$n_locid / p$nreg > 50) { 
    # reducing computation time of clusters by reducing the number of options
    p$sol_k_seq <- c(1:10, 12, 15, 20, 25, 35, 50, 70, 100, 150, 
                     seq(200, 1000, by = 100))
  } else {
    p$sol_k_seq <- NULL # use default settings
  }
  # cluster locations
  locid_sol_cl <- 
    cluster_locid(merra_solar, varname = p$sol_cf, 
                  locid_info = locid_sol_sf,
                  group = "reg_off", weight = "sol_MW_max",
                  # k = 1:p$sol_cl_max,
                  k = p$sol_k_seq,
                  max_loss = .01,
                  verbose = T)
  
  save(locid_sol_cl, file = fname); rm(fname)
} else {
  print(glue("Loading ", fname))
  load(fname); rm(fname)
}
```

### Aggregate

Aggregated capacity factors by region and cluster.

```{r agg_sol_cl, fig.height=8., warning=FALSE}
for (tol in p$sol_cl_tol) {
  # tol <- p$win_cl_tol[1]
  try({
  # create file identifier for clustered data
  sol_cl_mask <- rev_merra_cl_filemask(tol)
  
  # Select clusters based on tolerance parameter (see yaml header)

  # Select clusters based on parameters
  locid_sol_cl_k <- locid_sol_cl %>% 
    filter(sd_loss <= tol) %>% 
    group_by(reg_off) %>% 
    mutate(k_min = (k == min(k))) %>% ungroup() %>% 
    filter(k_min) %>% select(-k_min) %>%
    as.data.table()
  k_max <- max(locid_sol_cl_k$k) # for figures
  nreg_data <- length(unique(locid_sol_cl_k$reg_off)) # for figures
  # ToDo: split figures by ~20 groups for more than 20 regions
  rr <- unique(locid_sol_cl_k$reg_off)
  if (length(rr) > 20) {
    rr <- rr[1:20]
    print(paste0("Showing results for first 20 regions (out of ", 
          length(unique(locid_sol_cl_k$reg_off)), ")"))
  }
  # Cluster-loss figure
  locid_sol_cl_kk <- locid_sol_cl %>%
    group_by(reg_off, k) %>%
    summarise(sd_loss = max(sd_loss), N = max(N), .groups = "drop")
  
  locid_sol_cl_k_i <- locid_sol_cl_k %>%
    group_by(reg_off, k) %>%
    summarise(sd_loss = max(sd_loss), N = max(N), .groups = "drop")
  
  # a <- ggplot(locid_sol_cl_kk, aes(k, sd_loss)) +
  a <- ggplot(filter(locid_sol_cl_kk, reg_off %in% rr), aes(k, sd_loss)) +
    geom_line(color = "dodgerblue", linewidth = 1.5) +
    geom_point(color = "red", data = filter(locid_sol_cl_k_i, reg_off %in% rr)) +
    scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
    scale_x_continuous(breaks = rev_integer_breaks(5)) +
    labs(x = "Number of clusters (k)", 
         y = "loss, % of s.d.",
         title = "Aggregation lossess by number of clusters of solar sites") +
    facet_wrap(~paste0(reg_off, " (", N," locations)"), scales = "free",
               ncol = max(1, round(nrow(locid_sol_cl_k_i) / 5))) +
    rev_theme_map()
  
  if (tol < 1) {
    a <- a + 
      geom_hline(aes(yintercept = tol), col = "red", linetype = 2) +
      labs(subtitle = glue("Targeted tolerance ", 100 * tol, "% ",
                          "and minimal sufficient number of clusters",
                          " (red line and dot)"))
  } else {
    a <- a + 
      labs(subtitle = 
             glue("One cluster (sd_loss tolerance = ", 100 * tol, "%)"))
  }
  
  # suppressMessages({  
  print(a)
  fname <- glue("sol_cl_sd_loss_{sol_cl_mask}.png")
  ggsave(fname, a, 
         path = p$dir$fig_mod, 
         width = 6, height = 8, scale = 1.25)
  # })
  
  fname <- file.path(p$dir$data_mod, 
                     glue("locid_sol_cl_sf_{sol_cl_mask}.RData"))
  # suppressMessages({
  locid_sol_cl_sf <- locid_sol_sf %>%
    st_make_valid() %>%
    left_join(
      select(locid_sol_cl_k, any_of(c("reg_off", "region", "locid", "cluster")))
    ) %>%
    filter(!is.na(cluster)) %>%
    mutate(cluster = as.integer(cluster))
  # })
  print(glue("Writing ", fname))
  save(locid_sol_cl_sf, file = fname); rm(fname)

  a <- ggplot(locid_sol_cl_sf) +
    geom_sf(fill = "lightgrey", data = gis_sf)
  if (k_max > 10) {
    a <- a + geom_sf(aes(fill = as.integer(cluster)), color = NA) +
    scale_fill_viridis_c(option = "plasma", direction = 1, name = "Cluster")  
  } else {
    a <- a + geom_sf(aes(fill = as.factor(cluster)), color = NA) +
    scale_fill_viridis_d(option = "plasma", direction = 1, name = "Cluster")
  }
  a <- a +
  # a <- ggplot(locid_sol_cl_sf) +
  #   geom_sf(fill = "lightgrey", data = gis_sf) +
  #   geom_sf(aes(fill = cluster), color = NA) +
  #   scale_fill_viridis_c(option = "plasma", direction = 1, name = "Cluster") +
    geom_sf(color = alpha("black", 1), fill = NA, data = gis_sf) +
    labs(title = paste0("Clustered solar sites by region",
                        ", sd_loss <= ", tol * 100, "%")) +
    rev_theme_map()
  print(a)
  fname <- glue("locid_sol_cl_sf_{sol_cl_mask}.png")
  ggsave(fname, a, path = p$dir$fig_mod, 
         width = p$fig_map_width, height = p$fig_map_height, 
         scale = 1.25)

  # suppressMessages({
  merra_sol_cl <- merra_solar %>%
    filter(locid %in% unique(locid_sol_cl_sf$locid)) %>% # drop unused locations
    mutate(slice = timeslices::dtm2tsl(UTC)) %>% # create time-slice IDs
    select(locid, slice, starts_with("scf")) %>% # drop unused columns
    left_join( # merge with gis-data
      select(st_drop_geometry(locid_sol_cl_sf), 
             locid, region, reg_off, offshore, cluster, sol_MW_max), 
      relationship = "many-to-many"
    ) %>%
    group_by(slice, reg_off, cluster) %>%
    summarise(
      scf_fh = round(weighted.mean(scf_fh, sol_MW_max, na.rm = T), 3),
      scf_fl = round(weighted.mean(scf_fl, sol_MW_max, na.rm = T), 3),
      scf_tl = round(weighted.mean(scf_tl, sol_MW_max, na.rm = T), 3),
      scf_td = round(weighted.mean(scf_td, sol_MW_max, na.rm = T), 3),
      # geometry = st_union(geometry),
      .groups = "drop"
    ) %>%
    mutate(year = p$weather_year, .before = 1) %>%
    as.data.table()
  # })
  merra_sol_cl
  summary(merra_solar$scf_fl)
  summary(merra_solar$scf_tl)
  summary(merra_solar$scf_td)
  
  fname <- 
    file.path(p$dir$data_mod, glue("merra_sol_{sol_cl_mask}_{wyear}.fst"))
  print(glue("Writing ", fname))
  write_fst(merra_sol_cl, fname, 100); rm(fname)
  
  fname <- file.path(p$dir$data_mod, 
                     glue("sol_GW_max_sf_{sol_cl_mask}.RData"))
                     # glue("sol_GW_max_sf_{sol_cl_mask}_{wyear}.RData"))
  sol_GW_max_sf <- locid_sol_cl_sf %>% 
    # st_drop_geometry() %>% 
    # select(region, offshore, sol_MW_max) %>% 
    filter(sol_MW_max > 0) %>%
    st_make_valid() %>%
    group_by(region, reg_off, offshore, cluster) %>% 
    reframe(
      scf_fh = weighted.mean(scf_fh, sol_MW_max, na.rm = T),
      scf_fl = weighted.mean(scf_fl, sol_MW_max, na.rm = T),
      scf_tl = weighted.mean(scf_tl, sol_MW_max, na.rm = T),
      scf_td = weighted.mean(scf_td, sol_MW_max, na.rm = T),
      # area = sum(area, na.rm = T), 
      sol_GW_max = sum(sol_MW_max, na.rm = T) / 1e3, 
      geometry = rev_union_polygons(geometry)) %>%
    st_as_sf() %>% st_make_valid()
  print(glue("Writing: ", fname))
  save(sol_GW_max_sf, file = fname); rm(fname)
  
  a <- ggplot(sol_GW_max_sf) +
    geom_sf(fill = "lightgrey", data = gis_sf) +
    geom_sf(aes(fill = .data[[p$sol_cf]]), color = alpha("black", .5)) +
    scale_fill_viridis_c(option = "H", direction = 1) +
    labs(title = paste0("Clustered solar sites by region",
                        ", sd_loss <= ", tol * 100, "%")) +
    rev_theme_map()
  # suppressMessages({
    print(a)
    ggsave(glue("sol_GW_max_sf_{sol_cl_mask}_{wyear}.png"), a, 
           path = p$dir$fig_mod, width = p$fig_map_width, height = p$fig_map_height, 
           scale = 1.5)
  })
  # })
}
```

\newpage

## YAML parameters

A list of parameters from `r glue("config_{iso3c}.yml")` file YAML
header of the document, for reference.

```{r, echo=params$echo, include=params$print_parameters}
print(glue("Parameters: "))
str(p, give.head = F)

p_last <- p
save(p_last, file = file.path(p$dir$data_mod, "merra_last_config.RData"))
```

```{r, echo=params$echo, include=params$print_parameters}
# print(glue("Target outcome: ")
# cat(file.path(p$dir$data, ""))
```

```{r}
# installr::os.sleep()
```
